---
title: "shed-stats-all-days"
author: "Kyu Min Huh"
date: "2025-05-27"
output:
  html_document:
    code_folding: show
    fig_height: 5
    fig_width: 7
    toc: TRUE
    toc_depth: 5
    toc_float: 
      collapsed: false
      smooth_scroll: false
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = 'center', fig.pos = 'H')
```
                                    
# Synopsis 
In this analysis, I investigate physiological, behavioural, and social responses to a video hawk in wild blue tits and great tits at an automatic feeder. On each of 12 experimental days from December 2024 - February 2025, visiting PIT-tagged birds were assigned in 50:50 chance to control group, which never saw the predator, and treatment group, which were subjected to 1s of approaching goshawk video on their first visits and again on their subsequent visits in 50:50 chance. I compared bodyweight, initial leg acceleration (acc0), and three in-flight accelerations along the tunnel (acc1-3) on the day of experiment, a day before, and a day after ('day_since': -1 vs. 0 vs. 1).  

**Methods:**   

I used linear mixed model with date and individual ID as random effects and analysed blue tit and great tit separately as different variance was expected per species. All the final model data sets are balanced by experimental group ('control' vs. 'treatment) and the day visited respective to the experiment days ('day_since': -1 vs. 0 vs.1) per individual. This leaves blue tit with all 12 experimental days for the force data and 9 for the infrared; for great tit, 9 days for the force and 7 days for the infrared. 

**Key Results:**  

- bodyweight


Bodyweight increased by hour in both great tit (estimate = 0.06g) and blue tit (estimate = 0.02g) individuals. Blue Tit's bodyweight increased with rising temperature (0.01g/1 celsius).

For blue tit, treatment individuals increased bodyweight on the day (+0.11g) then a day after, both control and treatment group individuals decreased bodyweight (-0.13g) compared to a day before the experiment. 


Great tit females showed lower bodyweight than males, and 

In great tit, treatment birds on the day of experiment decreased bodyweight within individual (-0.36g, ~2% of bodyweight). In contrast, treatment blue tits showed higher bodyweight on the day within individual (estimate ~ 0.18g, 2% of bodyweight). However, the repeated number of hawk display showed cancelling effect to this increase (video display estimate = -0.14g, previous exposure = -0.02g). 

EDIT in the edited model Blue tit treatment individuals did increase bodyweight (0.11g), but control individuals decreased bodyweight (-0.16g) on the day of experiment. Both control and treatment group birds converged to a similar bodyweight a day after, which is -0.14g, which is around 2% decrease from two days ago. However, this change has not carried over to difference in leg and in-flight accelerations. 


- leg & in-flight accelerations

In both great tits and blue tits, the initial leg acceleration was lower on the day of the experiment, especially when the video hawk was shown ('scare_outcome' == 1) and decreased throughout the day by hour. In blue tit, the first in-flight acceleration, 0.165-m from the perch, was higher when video-hawk was displayed (est = +0.68 m/s^2). Great Tit showed no difference in in-flight acceleration between days nor experimental group, but showed a trend of 2+year birds faster at the first in-flight acceleration (acc1), then slower in further in-flight accelerations (acc2 & acc3) than the first year birds.

**Discussion:** 
Anti-predatory responses in wintering birds appear to be more immediate and flexible than expected in bodyweight, leg acceleration, and in-flight acceleration, as difference was only found on the day of experiment compared to a day before, and not the day after. In addition, the absence of main effect difference between control and treatment groups suggest the group-level social cues in face of predator attacks in mixed-flock birds. 

Blue tit demonstrates a diverging bodyweight strategy to great tit by increasing bodyweight at the presence of predator attack, but decreasing it by the frequency of exposure. This may be showcasing a complex strategy in face of both starvation and predation risk as a subdominant species, increasing bodyweight at overall presence of predation risk in preparation of losing a known food source, and then decreasing it only at repeated exposure to the predator to maintain the escape performance. 

Overall, video hawk display seems to have compromised initial leg acceleration in both species on the day of experiment. Blue tit appears to have compensated for the loss in leg acceleration with the increased first in-flight acceleration at 0.165 m. Great tit does not show this compensatory increase in any of its in-flight accelerations, but since no decrease was found neither, great tit may also have partially made up for its loss in leg acceleration with wings.  

# I. Data Carving

## libraries
```{r cars, echo = FALSE}
library(dplyr)
library(stringr)
library(readr)
library(fuzzyjoin)
library(tidyverse)
library(ggplot2)
library(lubridate)
library(ggpubr)
library(rstatix)
library(coin)
library(scales)
library(gridExtra)
library(ggthemes)
library(ggbreak)
library(lme4)
library(rptR)
library(reshape2)
library(olsrr)
library(gvlma)
library(performance)
library(PerformanceAnalytics)
library(usethis)
library(devtools)
library(redres)
library(EnvStats)
library(lmtest)
library(lmerTest)
library(AICcmodavg)
library(emmeans)
library(MuMIn)
library(sjPlot)
library(effects)
library(LaplacesDemon)
library(plm)
library(clubSandwich)
library(modelsummary)
library(marginaleffects)
library(DHARMa)
library(ggeffects)
library(robustlmm)
```



### 1. import dataset
```{r dataset}

# force data - day0
F_control_full <- read.delim(file ="C:\\Users\\kmh\\Documents\\shed-tits-data\\final_data\\F_control_full.csv", sep = ",") 

F_treats_full <- read.delim(file ="C:\\Users\\kmh\\Documents\\shed-tits-data\\final_data\\F_treats_full.csv", sep = ",") 

#day1
F_day1_full <- read.delim(file =  "C:\\Users\\kmh\\Documents\\shed-tits-data\\final_data\\F_day1_full.csv", sep = ",") 

#day-1 - being processed in VS code
F_dayminus1_full <- read.delim(file =  "C:\\Users\\kmh\\Documents\\shed-tits-data\\final_data\\F_dayminus1_full.csv", sep = ",") 

# IR data
control_ir_full <- read.delim(file = "C:\\Users\\kmh\\Documents\\shed-tits-data\\final_data\\ir_control_ordered.csv", sep = ",")

treats_ir_full <- read.delim(file = "C:\\Users\\kmh\\Documents\\shed-tits-data\\final_data\\ir_treats_ordered.csv", sep = ",") 

day1_ir_full <- read.delim(file = "C:\\Users\\kmh\\Documents\\shed-tits-data\\final_data\\day1_ir_full.csv", sep = ",")

dayminus1_ir_full <- read.delim(file = "C:\\Users\\kmh\\Documents\\shed-tits-data\\final_data\\day-1_ir_full.csv", sep = ",")

# temperature data 
temp_CP7 <- read.delim(file = "C:\\Users\\kmh\\Documents\\DATA\\2024-5\\temperature\\temp_CP7.csv", sep = ";", header = FALSE) %>%
  select(V2, V4) %>%
  slice(c(61966:73581),) #2024-11-06 00:00 - 2025-03-06 23:45
  

temp_CP28 <- read.delim(file = "C:\\Users\\kmh\\Documents\\DATA\\2024-5\\temperature\\temp_CP28.csv", sep = ";", header = FALSE) %>%
  select(V2, V4) %>%
  slice(c(62246:73861),)

colnames(temp_CP7) = c("datetime", "temperature_a")
colnames(temp_CP28) = c("datetime", "temperature_b")

temperature <- merge(temp_CP7, temp_CP28, by = "datetime")

  ## celsius column as a mean of two temperature measurements
temperature <- temperature %>%
  mutate(datetime = str_replace_all(datetime, "\\.", "-")) %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M")) %>% #aha so should not put %S as it's not in original formal %>%
  mutate(celsius = (temperature_a+temperature_b)/2)

plot(temperature$temperature_a, temperature$temperature_b)

```

### shape of datasets
```{r}

cat("F_control_full dim: ",dim(F_control_full)) #2063
cat("control_ir_full dim: ",dim(control_ir_full)) #1314
cat("F_treats_full dim: ",dim(F_treats_full)) #2215
cat("treats_ir_full dim: ",dim(treats_ir_full)) #1428

cat("F_day1_full dim: ",dim(F_day1_full)) #5808
cat("day1_ir_full dim: ",dim(day1_ir_full)) #5171
cat("F_day-1_full dim: ",dim(F_dayminus1_full)) #10465
cat("day-1_ir_full dim: ",dim(dayminus1_ir_full)) #6701


```

### 2. pooling control and treatment data 
```{r}
#F
F_pooled_full <-rbind(F_control_full, F_treats_full)

#IR
pooled_ir_full <-rbind(control_ir_full, treats_ir_full)

cat("F_pooled_full dim: ",dim(F_pooled_full)) #4278
cat("pooled_ir_full dim: ",dim(pooled_ir_full)) #2742
```

### 3. add hr column 
```{r}
#F
F_pooled_full$datetime <- as.POSIXct(F_pooled_full$datetime, format = "%Y-%m-%d %H:%M:%S")
F_day1_full$datetime <- as.POSIXct(F_day1_full$datetime, format = "%Y-%m-%d %H:%M:%S")
F_dayminus1_full$datetime <- as.POSIXct(F_dayminus1_full$datetime, format = "%Y-%m-%d %H:%M:%S")


F_pooled_full$hr <- hour(F_pooled_full$datetime)
F_day1_full$hr <- hour(F_day1_full$datetime)
F_dayminus1_full$hr <- hour(F_dayminus1_full$datetime)

#IR
pooled_ir_full$datetime <- as.POSIXct(pooled_ir_full$datetime, format = "%Y-%m-%d %H:%M:%S")
day1_ir_full$datetime <- as.POSIXct(day1_ir_full$datetime, format = "%Y-%m-%d %H:%M:%S")
dayminus1_ir_full$datetime <- as.POSIXct(dayminus1_ir_full$datetime, format = "%Y-%m-%d %H:%M:%S")

pooled_ir_full$hr <- hour(pooled_ir_full$datetime)
day1_ir_full$hr <- hour(day1_ir_full$datetime)
dayminus1_ir_full$hr <- hour(dayminus1_ir_full$datetime)

```

### 4. add temperature column
```{r}
temperature <- temperature[,!names(temperature) %in% c("temperature_a", "temperature_b")] 
```

```{r}
#merge by closest value in temperatre
F_pooled_full <- F_pooled_full %>%
  mutate(datetime_rounded = round_date(datetime, "15 minutes")) %>%
  mutate(datetime_rounded2 = datetime_rounded - hours(12)) %>%
  left_join(temperature, by = c("datetime_rounded" = "datetime")) %>%
  left_join(temperature, by = c("datetime_rounded2" = "datetime")) %>%
  rename( c(celsius.now = "celsius.x", celsius.12hr = "celsius.y"))

pooled_ir_full <- pooled_ir_full %>%
  mutate(datetime_rounded = round_date(datetime, "15 minutes")) %>%
  mutate(datetime_rounded2 = datetime_rounded - hours(12)) %>%
  left_join(temperature, by = c("datetime_rounded" = "datetime")) %>%
  left_join(temperature, by = c("datetime_rounded2" = "datetime")) %>%
  rename( c(celsius.now = "celsius.x", celsius.12hr = "celsius.y"))


F_day1_full <- F_day1_full %>%
  mutate(datetime_rounded = round_date(datetime, "15 minutes")) %>%
  mutate(datetime_rounded2 = datetime_rounded - hours(12)) %>%
  left_join(temperature, by = c("datetime_rounded" = "datetime")) %>%
  left_join(temperature, by = c("datetime_rounded2" = "datetime")) %>%
  rename( c(celsius.now = "celsius.x", celsius.12hr = "celsius.y"))

day1_ir_full <- day1_ir_full %>%
  mutate(datetime_rounded = round_date(datetime, "15 minutes")) %>%
  mutate(datetime_rounded2 = datetime_rounded - hours(12)) %>%
  left_join(temperature, by = c("datetime_rounded" = "datetime")) %>%
  left_join(temperature, by = c("datetime_rounded2" = "datetime")) %>%
  rename( c(celsius.now = "celsius.x", celsius.12hr = "celsius.y"))


F_dayminus1_full <- F_dayminus1_full %>%
  mutate(datetime_rounded = round_date(datetime, "15 minutes")) %>%
  mutate(datetime_rounded2 = datetime_rounded - hours(12)) %>%
  left_join(temperature, by = c("datetime_rounded" = "datetime")) %>%
  left_join(temperature, by = c("datetime_rounded2" = "datetime")) %>%
  rename( c(celsius.now = "celsius.x", celsius.12hr = "celsius.y"))

dayminus1_ir_full <- dayminus1_ir_full %>%
  mutate(datetime_rounded = round_date(datetime, "15 minutes")) %>%
  mutate(datetime_rounded2 = datetime_rounded - hours(12)) %>%
  left_join(temperature, by = c("datetime_rounded" = "datetime")) %>%
  left_join(temperature, by = c("datetime_rounded2" = "datetime")) %>%
  rename( c(celsius.now = "celsius.x", celsius.12hr = "celsius.y"))

```



### 5. bodyweight filter
```{r}

#bodyweight calculation
F_pooled_full$bodyweight = (F_pooled_full$bodyweight/9.8067) #bodyweight in kg
F_day1_full$bodyweight = (F_day1_full$bodyweight/9.8067)
F_dayminus1_full$bodyweight = (F_dayminus1_full$bodyweight/9.8067)

#bodyweight filter
#find rows where where 'bodyweight' > 0.25 OR 'species' == 'greti' & 'bodyweight' < 0.015 OR 'species' == 'bluti' & 'bodyweight > 0.15

bw_errors_day0 <- F_pooled_full %>%
  filter(
    bodyweight > 0.021 |
    (species == "greti" & bodyweight < 0.016) |
    (species == "bluti" & bodyweight > 0.013) |
    (species == "bluti" & bodyweight < 0.008)
  )

bw_errors_day1 <- F_day1_full %>%
  filter(
    bodyweight > 0.021 |
    (species == "greti" & bodyweight < 0.016) |
    (species == "bluti" & bodyweight > 0.013) |
    (species == "bluti" & bodyweight < 0.008)
  )

bw_errors_dayminus1 <- F_dayminus1_full %>%
  filter(
    bodyweight > 0.021 |
    (species == "greti" & bodyweight < 0.016) |
    (species == "bluti" & bodyweight > 0.013) |
    (species == "bluti" & bodyweight < 0.008)
  )

print(head(bw_errors_day0, 3))
print(head(bw_errors_day1, 3))
print(head(bw_errors_dayminus1, 3))

cat("Force day0 data before filtering: ",dim(F_pooled_full))

F_pooled_full <- setdiff(F_pooled_full, bw_errors_day0)
cat("Force day0 data after filtering: ",dim(F_pooled_full))

cat("Force day1: ",dim(F_day1_full))

F_day1_full <- setdiff(F_day1_full, bw_errors_day1)
cat("Force day1 after filtering: ",dim(F_day1_full))

cat("Force day-1: ",dim(F_dayminus1_full))
F_dayminus1_full <- setdiff(F_dayminus1_full, bw_errors_dayminus1)
cat("Force day-1 after filtering: ",dim(F_dayminus1_full))

```

### 6. int_acc and bodyweight adjustments
```{r}
#int_acc calculation
F_pooled_full$int_acc = (F_pooled_full$max_Ftot/F_pooled_full$bodyweight)
F_day1_full$int_acc = (F_day1_full$max_Ftot/F_day1_full$bodyweight)
F_dayminus1_full$int_acc = (F_dayminus1_full$max_Ftot/F_dayminus1_full$bodyweight)


# Drop rows where int_acc is NA
F_day0 <- F_pooled_full[!is.na(F_pooled_full$int_acc), ]
F_day1 <- F_day1_full[!is.na(F_day1_full$int_acc),]
F_dayminus1 <- F_dayminus1_full[!is.na(F_dayminus1_full$int_acc),]


dim(F_day0)#4509
dim(F_day1)#6315
dim(F_dayminus1)#9352
colnames(F_day0)
colnames(F_day1)
colnames(F_dayminus1)

#make bodyweight to g 
F_day0$bodyweight <- F_day0$bodyweight*1000
F_day1$bodyweight <- F_day1$bodyweight*1000
F_dayminus1$bodyweight <- F_dayminus1$bodyweight*1000


```


### 7. calculate angle
```{r}
#run only once
F_day0 <- F_day0 %>%
  mutate(takeoff_angle = atan(Fx_maxFtot/Fy_maxFtot)* (180/pi))

F_day1 <- F_day1 %>%
  mutate(takeoff_angle = atan(Fx_maxFtot/Fy_maxFtot)* (180/pi))

F_dayminus1 <- F_dayminus1 %>%
  mutate(takeoff_angle = atan(Fx_maxFtot/Fy_maxFtot)* (180/pi))

```




### 8. add 'day_since'column
```{r}

F_day0$day_since <- 0
F_day1$day_since <- 1
F_dayminus1$day_since <- -1

pooled_ir_full$day_since <- 0
day1_ir_full$day_since <- 1
dayminus1_ir_full$day_since <- -1 

colnames(F_day0)
colnames(F_day1)
colnames(F_dayminus1)



#for each 'event_no', only save the row with 'status' == 'Arrive' 
F_day1 <- F_day1 %>%
  group_by(event_no) %>%
  filter(status == "Arrive")

F_dayminus1 <- F_dayminus1 %>%
  group_by(event_no) %>%
  filter(status == "Arrive")

day1_ir_clean <- day1_ir_full %>%
  group_by(id, datetime) %>%
  filter(status == "Arrive")

dayminus1_ir_clean <- dayminus1_ir_full %>%
  group_by(id, datetime) %>%
  filter(status == "Arrive")

day0_ir <- pooled_ir_full
day1_ir <- day1_ir_clean
dayminus1_ir <- dayminus1_ir_clean

cat("F_day1 dim: ",dim(F_day1)) #3441
cat("F_day-1 dim: ",dim(F_dayminus1)) #5535
cat("day1_ir dim: ",dim(day1_ir))#2462
cat("day-1_ir dim: ",dim(dayminus1_ir)) #3194
```


### 9. add 'group' column
- tracking control & treatment birds on experimental days to the day after
#### a. group columns for F
```{r}
print(head(F_day0,3))

# Initialize an empty group_key dataframe
group_key_F <- data.frame(id = character(),
                        group = character(),
                        day = character())

# Process each day in F_scares
for (current_day in unique(F_day0$day)) {
  # Filter rows for the current day
  daily_data <- F_day0 %>%
    filter(day == current_day) %>%
    distinct(id, group, day)  # Find unique combinations of 'id', 'group', and 'day'
  #print(daily_data)
  
  # Append processed data to group_key dataframe
  group_key_F <- bind_rows(group_key_F, daily_data)
}

# View the resulting group_key dataframe
print(head(group_key_F,3))

```
#### b. group columns for IR
```{r}
#F_scares$day <- day(F_scares$visit_datetime)
print(head(day0_ir,3))

# Initialize an empty group_key dataframe
group_key_ir <- data.frame(id = character(),
                        group = character(),
                        day = character())

# Process each day in F_scares
for (current_day in unique(day0_ir$day)) {
  # Filter rows for the current day
  daily_data <- pooled_ir_full %>%
    filter(day == current_day) %>%
    distinct(id, group, day)  # Find unique combinations of 'id', 'group', and 'day'
  #print(daily_data)
  
  # Append processed data to group_key dataframe
  group_key_ir <- bind_rows(group_key_ir, daily_data)
}

# View the resulting group_key dataframe
print(head(group_key_ir,3))

```
#### c. Merge
```{r}
# F 
group_key_F$date_to_match_day0 = ymd(group_key_F$day)
group_key_F$date_to_match_day1 = date(group_key_F$date_to_match_day0) +1
group_key_F$date_to_match_dayminus1 = date(group_key_F$date_to_match_day0) -1

print(head(group_key_F,3))

## remove a 'day' column 
#F_day1 <- F_day1 %>% select(-day)
#F_dayminus1 <- F_dayminus1 %>% select(-day)

F_day1$date = date(F_day1$visit_datetime)
F_dayminus1$date = date(F_dayminus1$visit_datetime)

F_day1_groups <- merge(F_day1, group_key_F, by.x = c("id","date"), by.y = c("id","date_to_match_day1"), all.x = TRUE)
dim(F_day1_groups) #3024

F_dayminus1_groups <- merge(F_dayminus1, group_key_F, by.x = c("id","date"), by.y = c("id", "date_to_match_dayminus1"), all.x = TRUE)
# IR
group_key_ir$date_to_match_day0 = ymd(group_key_ir$day)
group_key_ir$date_to_match_day1 = date(group_key_ir$date_to_match_day0) +1
group_key_ir$date_to_match_dayminus1 = date(group_key_ir$date_to_match_day0) -1

print(head(group_key_ir,3))

## remove a 'day' column 
day1_ir <- day1_ir %>% select(-day)
dayminus1_ir <- dayminus1_ir %>% select(-day)

#make an operationable 'date' column 
day1_ir$date = date(day1_ir$visit_datetime)
dayminus1_ir$date = date(dayminus1_ir$visit_datetime)

#merge
day1_ir_groups <- merge(day1_ir, group_key_ir, by.x = c("id","date"), by.y = c("id","date_to_match_day1"), all.x = TRUE)

dayminus1_ir_groups <- merge(dayminus1_ir, group_key_ir, by.x = c("id","date"), by.y = c("id","date_to_match_dayminus1"), all.x = TRUE)

dim(day1_ir_groups) #2462
dim(dayminus1_ir_groups) #3194

```

#### d. Drop rows where visits are not assigned to any group
```{r}
# F 
cat("F_day1_groups with NAs: ",dim(F_day1_groups)) #3441
F_day1_groups_clean <- F_day1_groups[!is.na(F_day1_groups$group), ]
cat("F_day1_groups without NAs: ",dim(F_day1_groups_clean)) #2849. 

cat("F_day-1_groups with NAs: ",dim(F_dayminus1_groups)) #5535 
F_dayminus1_groups_clean <- F_dayminus1_groups[!is.na(F_dayminus1_groups$group), ]
cat("F_day-1_groups without NAs: ",dim(F_dayminus1_groups_clean)) #4535


# IR
cat("day1_ir_groups with NAs: ",dim(day1_ir_groups)) #2462
day1_ir_groups_clean <- day1_ir_groups[!is.na(day1_ir_groups$group), ]
cat("day1_ir_groups without NAs: ",dim(day1_ir_groups_clean)) #1902  not too bad! 

cat("day-1_ir_groups with NAs: ",dim(dayminus1_ir_groups)) #3194
dayminus1_ir_groups_clean <- dayminus1_ir_groups[!is.na(dayminus1_ir_groups$group), ]
cat("day-1_ir_groups without NAs: ",dim(dayminus1_ir_groups_clean)) #2521

```
#### e. IR - s_tot
```{r}

day0_ir <- day0_ir %>% 
  mutate(exit_s = (t12+t23+t34+t45)/1000)

day1_ir_groups_clean <- day1_ir_groups_clean %>% 
  mutate(exit_s = (t12+t23+t34+t45)/1000)

dayminus1_ir_groups_clean <- dayminus1_ir_groups_clean %>% 
  mutate(exit_s = (t12+t23+t34+t45)/1000)

```

#### e. binding into one dataset with day0 & day1
```{r}
#F
colnames(F_day1_groups_clean)
colnames(F_day0)
#max_Fx, max_Fy -> takeoff_angle
F_day1_final <- F_day1_groups_clean %>%select("id","visit_datetime","event_no","bodyweight","max_Ftot","Fx_maxFtot","Fy_maxFtot","takeoff_angle","species","sex","age","bw_filled","hr","int_acc","day_since","group","celsius.now", "celsius.12hr")

F_dayminus1_final <- F_dayminus1_groups_clean %>% select("id","visit_datetime","event_no","bodyweight","max_Ftot","Fx_maxFtot","Fy_maxFtot","takeoff_angle","species","sex","age","bw_filled","hr","int_acc","day_since","group","celsius.now", "celsius.12hr")

#adding scare_outcome == 0 for day1 
F_day1_final$scare_outcome <- 0 
F_day1_final$prev_scares <- 0 
F_dayminus1_final$scare_outcome <- 0 
F_dayminus1_final$prev_scares <- 0 

F_day0_final <- F_day0 %>% select("id","visit_datetime","event_no","bodyweight","max_Ftot","Fx_maxFtot","Fy_maxFtot","takeoff_angle","species","sex","age","bw_filled","hr","int_acc","day_since","group", "scare_outcome","prev_scares","celsius.now", "celsius.12hr") 

day01_F <- rbind(F_day0_final, F_day1_final)

day0minus_F <- rbind(F_dayminus1_final, F_day0_final)

all_days_F <- rbind(F_dayminus1_final, F_day0_final, F_day1_final)

cat("day01_F dim: ",dim(day01_F)) #7973
cat("day0minus_F dim: ",dim(day0minus_F)) #9659
cat("all_days_F dim: ",dim(all_days_F)) #12508


#IR 
colnames(day1_ir_groups_clean)
colnames(pooled_ir_full)

day1_ir_final <- day1_ir_groups_clean %>%select("id","visit_datetime","exit_s", "v1","v2","v3","v4","acc1","acc2","acc3","species","sex","age","hr","day_since","group","celsius.now", "celsius.12hr")

dayminus1_ir_final <- dayminus1_ir_groups_clean %>% select("id","visit_datetime","exit_s", "v1","v2","v3","v4","acc1","acc2","acc3","species","sex","age","hr","day_since","group","celsius.now", "celsius.12hr")
  
# adding scare_outcome for day1 and day-1
day1_ir_final$scare_outcome <- 0 
day1_ir_final$prev_scares <- 0 
dayminus1_ir_final$scare_outcome <- 0 
dayminus1_ir_final$prev_scares <- 0 

day0_ir_final <- day0_ir %>% select("id","visit_datetime","exit_s", "v1","v2","v3","v4","acc1","acc2","acc3","species","sex","age","hr","day_since","group", "scare_outcome", "prev_scares","celsius.now", "celsius.12hr") 

day01_ir <- rbind(day0_ir_final, day1_ir_final)
day0minus_ir <-rbind(dayminus1_ir_final, day0_ir_final)
all_days_ir <- rbind(dayminus1_ir_final, day0_ir_final, day1_ir_final)

# Drop rows with NA values and Inf in specified columns
day01_ir <- day01_ir %>%
  filter(!is.na(acc1) & !is.nan(acc1) & is.finite(acc1) &
         !is.na(acc2) & !is.nan(acc2) & is.finite(acc2) &
         !is.na(acc3) & !is.nan(acc3) & is.finite(acc3))

day0minus_ir <- day0minus_ir %>%
  filter(!is.na(acc1) & !is.nan(acc1) & is.finite(acc1) &
         !is.na(acc2) & !is.nan(acc2) & is.finite(acc2) &
         !is.na(acc3) & !is.nan(acc3) & is.finite(acc3))

all_days_ir <- all_days_ir %>%
  filter(!is.na(acc1) & !is.nan(acc1) & is.finite(acc1) &
         !is.na(acc2) & !is.nan(acc2) & is.finite(acc2) &
         !is.na(acc3) & !is.nan(acc3) & is.finite(acc3))

cat("day01_ir dim: ",dim(day01_ir)) #5322
cat("day0minus_ir dim: ",dim(day0minus_ir)) #5942
cat("all_days_ir dim: ",dim(all_days_ir)) #7843


```
#### f. 'date' column as categorical
```{r}
day01_F$visit_datetime <- as.POSIXct(day01_F$visit_datetime, format = "%Y-%m-%d %H:%M:%S")

day01_ir$visit_datetime <- as.POSIXct(day01_ir$visit_datetime, format = "%Y-%m-%d %H:%M:%S")

day0minus_F$visit_datetime <- as.POSIXct(day0minus_F$visit_datetime, format = "%Y-%m-%d %H:%M:%S")

day0minus_ir$visit_datetime <- as.POSIXct(day0minus_ir$visit_datetime, format = "%Y-%m-%d %H:%M:%S")

all_days_F$visit_datetime <- as.POSIXct(all_days_F$visit_datetime, format = "%Y-%m-%d %H:%M:%S")

all_days_ir$visit_datetime <- as.POSIXct(all_days_ir$visit_datetime, format = "%Y-%m-%d %H:%M:%S")

day01_F$date <- date(day01_F$visit_datetime)
day01_ir$date <- date(day01_ir$visit_datetime)
day01_F$month <- month(day01_F$visit_datetime)
day01_ir$month <- month(day01_ir$visit_datetime)

day0minus_F$date <- date(day0minus_F$visit_datetime)
day0minus_ir$date <- date(day0minus_ir$visit_datetime)
day0minus_F$month <- month(day0minus_F$visit_datetime)
day0minus_ir$month <- month(day0minus_ir$visit_datetime)

all_days_F$date <- date(all_days_F$visit_datetime)
all_days_ir$date <- date(all_days_ir$visit_datetime)
all_days_F$month <- month(all_days_F$visit_datetime)
all_days_ir$month <- month(all_days_ir$visit_datetime)
```

#### g. check scare_outcome
```{r}
day01_F <- day01_F %>%
  filter(!(group == "control" & scare_outcome == 1))

day01_ir <- day01_ir %>%
  filter(!(group == "control" & scare_outcome == 1 ))

day0minus_F <- day0minus_F %>%
  filter(!(group == "control" & scare_outcome == 1))

day0minus_ir <- day0minus_ir %>%
  filter(!(group == "control" & scare_outcome == 1 ))

all_days_F <- all_days_F %>%
  filter(!(group == "control" & scare_outcome == 1))

all_days_ir <- all_days_ir %>%
  filter(!(group == "control" & scare_outcome == 1 ))

dim(day01_F) #7322
dim(day0minus_F) #9655
dim(all_days_ir) #7839

dim(day01_ir) #5318
dim(day0minus_ir) #5938
dim(all_days_F) #11438

```
#### h. flight no. column
```{r}
all_days_F <- all_days_F %>%
  arrange(id, date, visit_datetime) %>%
  group_by(id, date) %>%
  mutate(flight_no = row_number())

all_days_ir <- all_days_ir %>%
  arrange(id, date, visit_datetime) %>%
  group_by(id, date) %>%
  mutate(flight_no = row_number())
  
```

# II. Model Assumption checks
## A->B?. Model selection
### 1. proposed linear model
```{r}
lmer_bw = lmer(bodyweight ~ 1 + day_since*group + scare_outcome + prev_scares + flight_no + sex + age + hr + date + celsius.now + (1|id), data = all_days_F)

lmer_maxFtot = lmer(max_Ftot ~ 1 + day_since*group + scare_outcome + prev_scares + flight_no + sex + age + hr + date + celsius.now + (1|id), data = all_days_F)

lmer_Fx = lmer(Fx_maxFtot ~ 1 + day_since*group + scare_outcome + prev_scares + flight_no + sex + age + hr + date + celsius.now + (1|id), data = all_days_F)

lmer_Fy = lmer(Fy_maxFtot ~ 1 + day_since*group + scare_outcome + prev_scares + flight_no + sex + age + hr + date + celsius.now + (1|id), data = all_days_F)

lmer0 = lmer(int_acc ~ 1 + day_since*group + scare_outcome + prev_scares + flight_no + sex + age + hr + date + celsius.now + (1|id), data =   all_days_F)

lmer_exits = lmer(exit_s ~ 1 + day_since*group + scare_outcome + prev_scares + flight_no + sex + age + hr + date + celsius.now + (1|id), data =   all_days_ir)


lmer_v1 = lmer(v1 ~ 1 + day_since*group + scare_outcome + prev_scares + flight_no + sex + age + hr + date + celsius.now + (1|id), data = all_days_ir)

lmer_v2 = lmer(v2 ~ 1 + day_since*group + scare_outcome + prev_scares + flight_no + sex + age + hr + date + celsius.now + (1|id), data = all_days_ir)

lmer_v3 = lmer(v3 ~ 1 + day_since*group + scare_outcome + prev_scares + flight_no + sex + age + hr + date + celsius.now + (1|id), data = all_days_ir)

lmer_v4 = lmer(v4 ~ 1 + day_since*group + scare_outcome + prev_scares + flight_no + sex + age + hr + date + celsius.now + (1|id), data = all_days_ir)



lmer1 = lmer(acc1 ~ 1 + day_since*group + scare_outcome + prev_scares + flight_no + sex + age + hr + date + celsius.now + (1|id), data =  all_days_ir)

lmer2 = lmer(acc2 ~  1 + day_since*group + scare_outcome + prev_scares + flight_no + sex + age + hr + date + celsius.now + (1|id), data = all_days_ir) 

lmer3 = lmer(acc3 ~  1 + day_since*group + scare_outcome + prev_scares + flight_no + sex + age + hr + date + celsius.now + (1|id), data = all_days_ir) 

```



### 2. outliers detection
#### a. outliers - forces
```{r}
test_bw <- rosnerTest(all_days_F$bodyweight, k=5) #no outliers
test_bw$all.stats

test_acc0 <- rosnerTest(all_days_F$int_acc, k=25)
test_acc0$all.stats

test_max_Ftot <- rosnerTest(all_days_F$max_Ftot, k=1)
test_max_Ftot$all.stats

test_Fx_maxFtot <- rosnerTest(all_days_F$Fx_maxFtot, k=1)
test_Fx_maxFtot$all.stats

test_Fy_maxFtot  <- rosnerTest(all_days_F$Fy_maxFtot, k=9)
test_Fy_maxFtot$all.stats

```
#### a. outliers - infrareds
```{r}


test_v1 <- rosnerTest(all_days_ir$v1, k = 5)
test_v1$all.stats

test_v2 <- rosnerTest(all_days_ir$v2, k = 9)
test_v2$all.stats

test_v4 <- rosnerTest(all_days_ir$v4, k = 8)
test_v4$all.stats

test_acc1 <- rosnerTest(all_days_ir$acc1, k=16)
test_acc1$all.stats 

test_acc2<- rosnerTest(all_days_ir$acc2, k=16)
test_acc2$all.stats 

test_acc3 <- rosnerTest(all_days_ir$acc3, k=68)
test_acc3$all.stats

```

```{r}
test_exits <- rosnerTest(all_days_ir$exit_s, k = 68)
#more
test_exits$all.stats

test_v3 <- rosnerTest(all_days_ir$v3, k = 68)
test_v3$all.stats #more
```

#### b.1. collecting outlier row numbers
```{R}
#no outliers in bodyweight
#F
out_acc0 <- test_acc0$all.stats %>%
  filter(Outlier == "TRUE") %>%
  select(Obs.Num)

out_max_Ftot <- test_max_Ftot$all.stats %>%
  filter(Outlier == "TRUE") %>%
  select(Obs.Num)

out_Fx <- test_Fx_maxFtot$all.stats %>%
  filter(Outlier == "TRUE") %>%
  select(Obs.Num)

out_Fy <- test_Fy_maxFtot$all.stats %>%
  filter(Outlier == "TRUE") %>%
  select(Obs.Num)



#IR
out_exits <- test_exits$all.stats %>%
  filter(Outlier == "TRUE") %>%
  select(Obs.Num)

out_v1 <- test_v1$all.stats %>%
  filter(Outlier == "TRUE") %>%
  select(Obs.Num)

out_v2 <- test_v2$all.stats %>%
  filter(Outlier == "TRUE") %>%
  select(Obs.Num)

out_v3 <- test_v3$all.stats %>%
  filter(Outlier == "TRUE") %>%
  select(Obs.Num)

out_v4 <- test_v4$all.stats %>%
  filter(Outlier == "TRUE") %>%
  select(Obs.Num)

out_acc1 <- test_acc1$all.stats %>%
  filter(Outlier == "TRUE") %>%
  select(Obs.Num)

out_acc2 <- test_acc2$all.stats %>%
  filter(Outlier == "TRUE") %>%
  select(Obs.Num)

out_acc3 <- test_acc3$all.stats %>%
  filter(Outlier == "TRUE") %>%
  select(Obs.Num)

outliers_F <-rbind(out_acc0, out_max_Ftot, out_Fx, out_Fy)
outliers_IR <- rbind(out_exits, out_v1, out_v2, out_v3, out_v4, out_acc1, out_acc2, out_acc3) %>%
  distinct
dim(outliers_F) #36 outliers
dim(outliers_IR)
 #194 outliers

```


#### c. datasets without outliers
```{r}
force.out <- all_days_F[-c(outliers_F$Obs.Num),]
ir.out <- all_days_ir[-c(outliers_IR$Obs.Num),]
dim(ir.out) #7645
dim(force.out) #11408
```




## B ->A?. data polishing
### 4. dataset balancing
```{r}
#date
force.out$date <- as.character(force.out$date)
ir.out$date <- as.character(ir.out$date)

# Define mapping for date labels
date_map <- c("2024-12-11" = "Dec12", "2024-12-12" = "Dec12", "2024-12-13" = "Dec12","2024-12-14" = "Dec15", "2024-12-15" = "Dec15", "2024-12-16" = "Dec15","2024-12-22" = "Dec23", "2024-12-23" = "Dec23", "2024-12-24" = "Dec23","2024-12-30" = "Dec31", "2024-12-31" = "Dec31", "2025-01-01" = "Dec31","2025-01-02" = "Jan3", "2025-01-03" = "Jan3", "2025-01-04" = "Jan3","2025-01-20" = "Jan21", "2025-01-21" = "Jan21", "2025-01-22" = "Jan21","2025-01-23" = "Jan24", "2025-01-24" = "Jan24", "2025-01-25" = "Jan24","2025-01-29" = "Jan30", "2025-01-30" = "Jan30", "2025-01-31" = "Jan30","2025-02-05" = "Feb6", "2025-02-06" = "Feb6", "2025-02-07" = "Feb6","2025-02-11" = "Feb12", "2025-02-12" = "Feb12", "2025-02-13" = "Feb12","2025-02-15" = "Feb16", "2025-02-16" = "Feb16", "2025-02-17" = "Feb16","2025-02-18" = "Feb19", "2025-02-19" = "Feb19", "2025-02-20" = "Feb19")

# Assign labels to the dataframe using the mapping
force.out$date_label <- date_map[as.character(force.out$date)]
ir.out$date_label <- date_map[as.character(ir.out$date)]


#subset by species
force.out.greti <- subset(force.out, species == "greti") 


force.out.bluti <- subset(force.out, species == "bluti")

ir.out.greti <- subset(ir.out, species == "greti")
ir.out.bluti <- subset(ir.out, species == "bluti")
```
```{r}
#looking if good amount of individuals are present for all days 
count_table_greti <- force.out.greti %>%
  group_by(id, day_since) %>%
  summarise(count = n(), .groups = "drop") %>%
  tidyr::pivot_wider(names_from = day_since, values_from = count, values_fill = list(count = 0))

print(count_table_greti)

count_table_bluti <- force.out.bluti %>%
  group_by(id, day_since) %>%
  summarise(count = n(), .groups = "drop") %>%
  tidyr::pivot_wider(names_from = day_since, values_from = count, values_fill = list(count = 0))

print(count_table_bluti)

#dataset balancing 

# enforced a bit stronger criteria as 14g might be two blue tit or 1.5 blue tits. 
force.out.bluti_bal <- force.out.bluti %>% filter(int_acc <50) %>% filter(bodyweight > 8 & bodyweight < 13) %>% group_by(id) %>%
filter(all(c(-1, 0, 1) %in% day_since)) %>% na.omit() %>% ungroup()

ir.out.bluti_bal <- ir.out.bluti %>% filter(!date_label %in% c("Feb16","Feb19","Jan30")) %>% group_by(id) %>%
filter(all(c(-1, 0, 1) %in% day_since)) %>% na.omit() %>% ungroup()

#the dates where there is only one data point or not balanced
force.out.greti_bal <- force.out.greti %>% filter(!date_label %in% c("Dec23","Jan21","Jan30")) %>% group_by(id) %>%
filter(all(c(-1, 0, 1) %in% day_since)) %>% na.omit() %>% ungroup()

ir.out.greti_bal <- ir.out.greti %>% filter(!date_label %in% c("Dec23","Feb16","Feb19","Jan21","Jan30")) %>%group_by(id) %>%
filter(all(c(-1, 0, 1) %in% day_since)) %>% na.omit() %>% ungroup()

dim(force.out.bluti)
dim(force.out.bluti_bal) #3,126 flights out, 64% 

dim(ir.out.bluti)
dim(ir.out.bluti_bal) #1,455 flights out, 72%

dim(force.out.greti)
dim(force.out.greti_bal) #908 flights out, 76%

dim(ir.out.greti)
dim(ir.out.greti_bal) #957 flights out, 63%

```
```{r}

force.out.bluti_bal_table <- force.out.bluti_bal %>%
  group_by(id) %>%
  summarise(visits = n()) %>%
  summarise(
    indiv.no = n(),
    flight.total = sum(visits),
    mean.visits = mean(visits),
    sd.visits = sd(visits)
  )

force.out.greti_bal_table <- force.out.greti_bal %>%
  group_by(id) %>%
  summarise(visits = n()) %>%
  summarise(
    indiv.no = n(),
    flight.total = sum(visits),
    mean.visits = mean(visits),
    sd.visits = sd(visits)
  )

ir.out.bluti_bal_table <- ir.out.bluti_bal %>%
  group_by(id) %>%
  summarise(visits = n()) %>%
  summarise(
    indiv.no = n(),
    flight.total = sum(visits),
    mean.visits = mean(visits),
    sd.visits = sd(visits)
  )

ir.out.greti_bal_table <- ir.out.greti_bal %>%
  group_by(id) %>%
  summarise(visits = n()) %>%
  summarise(
    indiv.no = n(),
    flight.total = sum(visits),
    mean.visits = mean(visits),
    sd.visits = sd(visits)
  )

force.out.bluti_bal_table
force.out.greti_bal_table
ir.out.bluti_bal_table
ir.out.greti_bal_table
```
#### indiv visiting number
```{r}
indiv_visit_no <- force.out.bluti_bal %>%
  count(date_label, id, day_since, group) %>%
  pivot_wider(names_from = day_since, values_from = n, values_fill = list(n = 0))
colnames(indiv_visit_no)[4:6] <- c("day-1", "day0", "day1")


indiv_visit_no_more <-indiv_visit_no %>%
  filter(`day-1` >10 & day0 >10 & day1 >10)


# Randomly pick 5 rows
random_rows <- indiv_visit_no_more %>% sample_n(5)
random_rows
```
#### number of visiting dates per individual
```{r}

dates_visited <- force.out.bluti_bal %>%
  group_by(id) %>%
  summarise(unique_dates = n_distinct(date)) %>%
  summarise(
    mean_unique_dates = mean(unique_dates),
    sd_unique_dates = sd(unique_dates)
  )

dates_visited
```

### 5. checking correlations
#### 5.a. na checks
```{r}
na_counts <- force.out.bluti_bal %>%
     select_if(is.numeric) %>%
      summarise_all(~ sum(is.na(.)))
na_counts
```

#### 5.b. correlation checks (can skip)
- understanbly high correlation between celsius.now x celsius.12 (12 hours before) and hr x flight_no but no one dangerously high enough to omit
- 
```{r}

#need 'id', 'exp_order', 'hr', 'bodyweight'
usr <- par("usr") 
par(usr=usr)


force.out.bluti_bal$prev_scares <-as.numeric(force.out.bluti_bal$prev_scares)
force.out.bluti_bal$flight_no <- as.numeric(force.out.bluti_bal$flight_no)

ir.out.bluti_bal$prev_scares <-as.numeric(ir.out.bluti_bal$prev_scares)
ir.out.bluti_bal$flight_no <- as.numeric(ir.out.bluti_bal$flight_no)

force.out.greti_bal$prev_scares <-as.numeric(force.out.greti_bal$prev_scares)
force.out.greti_bal$flight_no <- as.numeric(force.out.greti_bal$flight_no)

ir.out.greti_bal$prev_scares <-as.numeric(ir.out.greti_bal$prev_scares)
ir.out.greti_bal$flight_no <- as.numeric(ir.out.greti_bal$flight_no)


#for some reason 'celsius' not charting 
force.out.bluti_bal_corcheck <- force.out.bluti_bal %>% select( 'day_since','bodyweight','scare_outcome','celsius.now', 'celsius.12hr','hr', 'prev_scares', 'flight_no')
cor.chart_bluti_F = chart.Correlation(force.out.bluti_bal_corcheck, histogram=TRUE, pch=19)



ir.out.bluti_bal_corcheck <- ir.out.bluti_bal %>% select( 'day_since','scare_outcome','celsius.now', 'celsius.12hr','hr', 'prev_scares','flight_no')
cor.chart_bluti_ir = chart.Correlation(ir.out.bluti_bal_corcheck, histogram=TRUE, pch=19)




force.out.greti_bal_corcheck <- force.out.greti_bal %>% select( 'day_since','bodyweight','scare_outcome','celsius.now', 'celsius.12hr','hr', 'prev_scares', 'flight_no' )
cor.chart_greti_F = chart.Correlation(force.out.greti_bal_corcheck, histogram=TRUE, pch=19)


ir.out.greti_bal_corcheck <- ir.out.greti_bal %>% select( 'day_since','scare_outcome','celsius.now', 'celsius.12hr','hr', 'prev_scares','flight_no' )
cor.chart_greti_ir = chart.Correlation(ir.out.greti_bal_corcheck, histogram=TRUE, pch=19)

```

### 6. datasets typing
```{r}

#scare_outcome (to character)
force.out.bluti_bal$scare_outcome <- as.character(force.out.bluti_bal$scare_outcome)
ir.out.bluti_bal$scare_outcome <- as.character(ir.out.bluti_bal$scare_outcome)

force.out.greti_bal$scare_outcome <- as.character(force.out.greti_bal$scare_outcome)
ir.out.greti_bal$scare_outcome <- as.character(ir.out.greti_bal$scare_outcome)

#age (to character)
force.out.bluti_bal$age <- as.character(force.out.bluti_bal$age)
ir.out.bluti_bal$age <- as.character(ir.out.bluti_bal$age)

force.out.greti_bal$age <- as.character(force.out.greti_bal$age)
ir.out.greti_bal$age <- as.character(ir.out.greti_bal$age)

#month (to character)
force.out.bluti_bal$month <- as.character(force.out.bluti_bal$month)
ir.out.bluti_bal$month <- as.character(ir.out.bluti_bal$month)

force.out.greti_bal$month <- as.character(force.out.greti_bal$month)
ir.out.greti_bal$month <- as.character(ir.out.greti_bal$month)

#day_since (to character)
force.out.bluti_bal$day_since <- as.character(force.out.bluti_bal$day_since)
ir.out.bluti_bal$day_since <- as.character(ir.out.bluti_bal$day_since)

force.out.greti_bal$day_since <- as.character(force.out.greti_bal$day_since)
ir.out.greti_bal$day_since <- as.character(ir.out.greti_bal$day_since)

# making group -> exp_group so that marginal effect packages are happy 
force.out.greti_bal <- force.out.greti_bal %>%
  rename(exp_group = group)
force.out.bluti_bal <- force.out.bluti_bal %>%
  rename(exp_group = group)

ir.out.greti_bal<- ir.out.greti_bal %>%
  rename(exp_group = group)
ir.out.bluti_bal<- ir.out.bluti_bal %>%
  rename(exp_group = group)

# changing 'scare_outcome' 


force.out.bluti_bal$scare_outcome <- factor(force.out.bluti_bal$scare_outcome, levels = c(0, 1), labels = c("normal", "escape"))

force.out.greti_bal$scare_outcome <- factor(force.out.greti_bal$scare_outcome, levels = c(0, 1), labels = c("normal", "escape"))

ir.out.bluti_bal$scare_outcome <- factor(ir.out.bluti_bal$scare_outcome, levels = c(0, 1), labels = c("normal", "escape"))

ir.out.greti_bal$scare_outcome <- factor(ir.out.greti_bal$scare_outcome, levels = c(0, 1), labels = c("normal", "escape"))




```


## C. Check assumption graphs 
### BLUTI {.tabset} 

### *control vs. treatment*
#### bodyweight
##### model assumption
```{r}

lmer_bw_bluti_group = lmer(bodyweight ~ 1  + day_since*exp_group + scare_outcome + prev_scares + flight_no  + sex + age + hr+ celsius.now + date_label + (1|id), data = force.out.bluti_bal)

check_collinearity(lmer_bw_bluti_group)
#celsius.now + celsius.12hr + month

# Computes the default residuals (raw conditional)
raw_cond_bw <- compute_redres(lmer_bw_bluti_group)

#linearity - can't test as all the terms are categorical 
hist(raw_cond_bw) #normality in histogram
plot_redres(lmer_bw_bluti_group) #heteroskedasticity
plot_resqq(lmer_bw_bluti_group) #normality of residuals - heavy tailed and bimodal?
plot_ranef(lmer_bw_bluti_group) #normality of random effects residuals

simulation_output <- simulateResiduals(fittedModel = lmer_bw_bluti_group, n = 1000)
plot(simulation_output)
```

##### robustlmm {.tab}
```{r}

lmer_bw_bluti_group = lmer(bodyweight ~ 1 + day_since*exp_group + scare_outcome + prev_scares + flight_no + age + sex + hr+ celsius.now + date_label+(1|id), data = force.out.bluti_bal)

rlmer_bw_bluti_group = rlmer(bodyweight ~ 1 + day_since*exp_group + scare_outcome + prev_scares + flight_no + age + sex + hr+ celsius.now + date_label+(1|id), data = force.out.bluti_bal)

summary(rlmer_bw_bluti_group)
plot(rlmer_bw_bluti_group)
```

```{r}
rlmer2_bw_group <- update(rlmer_bw_bluti_group, rho.sigma.e = psi2propII(smoothPsi, k = 1.66), 
                 rho.sigma.b = psi2propII(smoothPsi, k = 1.66)) #both fixed and random effects -> k = 1.66 
```

```{r}
rlmer3_bw_group <- update(rlmer2_bw_group,  rho.sigma.e = psi2propII(smoothPsi, k = 2.28), 
                 rho.sigma.b = psi2propII(smoothPsi, k = 2.28))
```

```{r}
compare(lmer_bw_bluti_group, rlmer_bw_bluti_group, rlmer2_bw_group, rlmer3_bw_group, show.rho.functions = FALSE)
#0.497, 0.344, 0.347, 0.382. just rlmer
```
###### tab_model
```{r}
tab_model(rlmer_bw_bluti_group)
```
###### plot 
```{r}

#scare response
rlmer_bw_scare_predict <- predict_response(rlmer_bw_bluti_group, "scare_outcome", margin = "empirical")


bw_scare <- plot(rlmer_bw_scare_predict)+ labs(x ="", y = "bodyweight (g)", title = "") 
bw_scare

#group response
rlmer_bw_bluti_predict <- predict_response(rlmer_bw_bluti_group, c("day_since","exp_group"),margin = "empirical")
bw_group <- plot(rlmer_bw_bluti_predict) + labs(x = "", y = "bodyweight (g)", title = "", colour = "group") +scale_color_manual(values = c("#0072B2", "#CC79A7"))+theme(legend.postion = "none")

bw_group 

rlmer_bw_bluti_interact <- predict_response(rlmer_bw_bluti_group, c("day_since","exp_group"),margin = "marginalmeans")
```
###### test
```{r}
emm_options(rg.limit = 40000)

test_predictions(rlmer_bw_scare_predict, collapse_levels = TRUE)
test_predictions(rlmer_bw_bluti_predict, collapse_levels = TRUE, p_adjust = "sidak")

test_predictions(rlmer_bw_bluti_interact, test = "interaction", p_adjust = "sidak")
# no 
```

#### peak_force
##### model assumption
```{r}

lmer_maxF_bluti_group = lmer(max_Ftot ~ 1  + day_since*exp_group + scare_outcome + prev_scares + flight_no  + sex + age + hr+ celsius.now + date_label + (1|id), data = force.out.bluti_bal)


#celsius.now + celsius.12hr + month

# Computes the default residuals (raw conditional)
raw_cond_maxF <- compute_redres(lmer_maxF_bluti_group)

#linearity - can't test as all the terms are categorical 
hist(raw_cond_maxF) #normality in histogram
plot_redres(lmer_maxF_bluti_group) #heteroskedasticity
plot_resqq(lmer_maxF_bluti_group) #normality of residuals - heavy tailed and bimodal?
plot_ranef(lmer_maxF_bluti_group) #normality of random effects residuals

simulation_output <- simulateResiduals(fittedModel = lmer_maxF_bluti_group, n = 1000)
plot(simulation_output)
```

##### robustlmm {.tab}
```{r}

lmer_maxF_bluti_group = lmer(max_Ftot ~ 1 + day_since*exp_group + scare_outcome + prev_scares + flight_no + age + sex + hr+ celsius.now + date_label+(1|id), data = force.out.bluti_bal)

rlmer_maxF_bluti_group = rlmer(max_Ftot ~ 1 + day_since*exp_group + scare_outcome + prev_scares + flight_no + age + sex + hr+ celsius.now + date_label+(1|id), data = force.out.bluti_bal)

summary(rlmer_maxF_bluti_group)
plot(rlmer_maxF_bluti_group)
```

```{r}
rlmer2_maxF_group <- update(rlmer_maxF_bluti_group, rho.sigma.e = psi2propII(smoothPsi, k = 1.66), 
                 rho.sigma.b = psi2propII(smoothPsi, k = 1.66)) #both fixed and random effects -> k = 1.66 
```

```{r}
rlmer3_maxF_group <- update(rlmer2_maxF_group,  rho.sigma.e = psi2propII(smoothPsi, k = 2.28), 
                 rho.sigma.b = psi2propII(smoothPsi, k = 2.28))
```

```{r}
compare(lmer_maxF_bluti_group, rlmer_maxF_bluti_group, rlmer2_maxF_group, rlmer3_maxF_group, show.rho.functions = FALSE)
#0.125, 0.133, 0.129, 0.127. 
```
###### tab_model
```{r}
tab_model(rlmer3_maxF_group)
```
###### plot 
```{r}
#scare_outcome
rlmer_maxF_scare_predict <- predict_response(rlmer3_maxF_group, "scare_outcome", margin = "empirical")
maxF_scare <- plot(rlmer_maxF_scare_predict) + labs(x ="", y = "max F (N)", title = "") 

maxF_scare

#within and between groups
rlmer_maxF_group_predict <- predict_response(rlmer3_maxF_group, c("day_since","exp_group"),margin = "empirical")

maxF_group <- plot(rlmer_maxF_group_predict) + labs(x = "", y = "max F (N)", title = "", colour = "group") +scale_color_manual(values = c("#0072B2", "#CC79A7"))
maxF_group

rlmer_maxF_group_interact <- predict_response(rlmer3_maxF_group, c("day_since","exp_group"), margin = "marginalmeans")
```
###### test
```{r}
emm_options(rg.limit = 40000)

test_predictions(rlmer_maxF_scare_predict)

test_predictions(rlmer_maxF_group_predict, collapse_levels = TRUE, p_adjust = "sidak")
#within-group difference on the experimental day for both control & treatment groups 

test_predictions(rlmer_maxF_group_interact, test = "interaction", p_adjust = "sidak") 
# there is a difference between treatment vs. control group on day 0. 
```
#### Fx
##### model assumption
```{r}

lmer_Fx_bluti_group = lmer(Fx_maxFtot ~ 1  + day_since*exp_group + scare_outcome + prev_scares + flight_no  + sex + age + hr+ celsius.now + date_label + (1|id), data = force.out.bluti_bal)


#celsius.now + celsius.12hr + month

# Computes the default residuals (raw conditional)
raw_cond_Fx <- compute_redres(lmer_Fx_bluti_group)

#linearity - can't test as all the terms are categorical 
hist(raw_cond_Fx) #normality in histogram
plot_redres(lmer_Fx_bluti_group) #heteroskedasticity
plot_resqq(lmer_Fx_bluti_group) #normality of residuals - heavy tailed and bimodal?
plot_ranef(lmer_Fx_bluti_group) #normality of random effects residuals

simulation_output <- simulateResiduals(fittedModel = lmer_maxF_bluti_group, n = 1000)
plot(simulation_output)
```

##### robustlmm {.tab}
```{r}

lmer_Fx_bluti_group = lmer(Fx_maxFtot ~ 1 + day_since*exp_group + scare_outcome + prev_scares + flight_no + age + sex + hr+ celsius.now + date_label+(1|id), data = force.out.bluti_bal)

rlmer_Fx_bluti_group = rlmer(Fx_maxFtot ~ 1 + day_since*exp_group + scare_outcome + prev_scares + flight_no + age + sex + hr+ celsius.now + date_label+(1|id), data = force.out.bluti_bal)

summary(rlmer_Fx_bluti_group)
plot(rlmer_Fx_bluti_group)
```

```{r}
rlmer2_Fx_group <- update(rlmer_Fx_bluti_group, rho.sigma.e = psi2propII(smoothPsi, k = 1.66), 
                 rho.sigma.b = psi2propII(smoothPsi, k = 1.66)) #both fixed and random effects -> k = 1.66 
#warning
```

```{r}
rlmer3_Fx_group <- update(rlmer2_Fx_group,  rho.sigma.e = psi2propII(smoothPsi, k = 2.28), 
                 rho.sigma.b = psi2propII(smoothPsi, k = 2.28))
```

```{r}
compare(lmer_Fx_bluti_group, rlmer_Fx_bluti_group, rlmer2_Fx_group, rlmer3_Fx_group, show.rho.functions = FALSE)
#0.115, 0.0994, 0.0995, 0.105. just rlmer
```
###### tab_model
```{r}
tab_model(rlmer_Fx_bluti_group)
```
###### plot 
```{r}
# scare comparison 
rlmer_Fx_scare_predict <- predict_response(rlmer_Fx_bluti_group, "scare_outcome",margin = "empirical")
Fx_scare <- plot(rlmer_Fx_scare_predict) + labs(x ="", y = "Fx (N)", title = "")

Fx_scare

# group comparison
rlmer_Fx_group_predict <- predict_response(rlmer_Fx_bluti_group, c("day_since","exp_group"),margin = "empirical")
Fx_group <- plot(rlmer_Fx_group_predict)+ labs(x = "", y = "Fx (N)", title = "", colour = "group") +scale_color_manual(values = c("#0072B2", "#CC79A7"))  
Fx_group

rlmer_Fx_bluti_interact <- predict_response(rlmer_Fx_bluti_group, c("day_since","exp_group"),margin = "marginalmeans")
```
###### test
```{r}
emm_options(rg.limit = 40000)

test_predictions(rlmer_Fx_scare_predict)


test_predictions(rlmer_Fx_group_predict, collapse_levels = TRUE, p_adjust = "sidak")
#within-individual significantly different on the experimental day. 

test_predictions(rlmer_Fx_bluti_interact, test = "interaction", p_adjust = "sidak") 
# control vs. treatment significantly different on the day. 
```
#### Fy
##### model assumption
```{r}

lmer_Fy_bluti_group = lmer(-Fy_maxFtot ~ 1  + day_since*exp_group + scare_outcome + prev_scares + flight_no  + sex + age + hr+ celsius.now + date_label + (1|id), data = force.out.bluti_bal)


#celsius.now + celsius.12hr + month

# Computes the default residuals (raw conditional)
raw_cond_Fy <- compute_redres(lmer_Fy_bluti_group)

#linearity - can't test as all the terms are categorical 
hist(raw_cond_Fy) #normality in histogram
plot_redres(lmer_Fy_bluti_group) #heteroskedasticity
plot_resqq(lmer_Fy_bluti_group) #normality of residuals - heavy tailed and bimodal?
plot_ranef(lmer_Fy_bluti_group) #normality of random effects residuals

simulation_output <- simulateResiduals(fittedModel = lmer_Fy_bluti_group, n = 1000)
plot(simulation_output)
```

##### robustlmm {.tab}
```{r}

lmer_Fy_bluti_group = lmer(-Fy_maxFtot ~ 1 + day_since*exp_group + scare_outcome + prev_scares + flight_no + age + sex + hr+ celsius.now + date_label+(1|id), data = force.out.bluti_bal)

rlmer_Fy_bluti_group = rlmer(-Fy_maxFtot ~ 1 + day_since*exp_group + scare_outcome + prev_scares + flight_no + age + sex + hr+ celsius.now + date_label+(1|id), data = force.out.bluti_bal)

summary(rlmer_Fy_bluti_group)
plot(rlmer_Fy_bluti_group)
```

```{r}
rlmer2_Fy_group <- update(rlmer_Fy_bluti_group, rho.sigma.e = psi2propII(smoothPsi, k = 1.66), 
                 rho.sigma.b = psi2propII(smoothPsi, k = 1.66)) #both fixed and random effects -> k = 1.66 
```

```{r}
rlmer3_Fy_group <- update(rlmer2_Fy_group,  rho.sigma.e = psi2propII(smoothPsi, k = 2.28), 
                 rho.sigma.b = psi2propII(smoothPsi, k = 2.28))
```

```{r}
compare(lmer_Fy_bluti_group, rlmer_Fy_bluti_group, rlmer2_Fy_group, rlmer3_Fy_group, show.rho.functions = FALSE)
#0.129, 0.131, 0.128, 0.129 rlmer2
```
###### tab_model
```{r}
tab_model(rlmer2_Fy_group)
```
###### plot 
```{r}
#scare_outcome
rlmer_Fy_scare_predict <- predict_response(rlmer2_Fy_group, "scare_outcome",margin = "empirical")
Fy_scare <- plot(rlmer_Fy_scare_predict) + labs(x ="", y = "Fy (N)", title = "")

Fy_scare

#group
rlmer_Fy_group_predict <- predict_response(rlmer2_Fy_group, c("day_since","exp_group"),margin = "empirical")
Fy_group <- plot(rlmer_Fy_group_predict) + labs(x = "", y = "Fy (N)", title = "", colour = "group") +scale_color_manual(values = c("#0072B2", "#CC79A7"))   
Fy_group

#interaction
rlmer_Fy_bluti_interact <- predict_response(rlmer2_Fy_group, c("day_since","exp_group"),margin = "marginalmeans")
```
###### test
```{r}
emm_options(rg.limit = 40000)

#scare_outcome
test_predictions(rlmer_Fy_scare_predict)

#group
test_predictions(rlmer_Fy_group_predict, collapse_levels = TRUE, p_adjust = "sidak")
#within-group significant on day 0

test_predictions(rlmer_Fy_bluti_interact, test = "interaction", p_adjust = "sidak") 
# -1 -> 0 significant (0.024). 0 to 1 less significant (p = 0.055)
```
#### int_acc (acc0)
##### model assumption
```{r}

lmer_acc0_bluti_group = lmer(int_acc ~ 1  + day_since*exp_group + scare_outcome + prev_scares + flight_no  + sex + age + hr+ celsius.now + date_label + (1|id), data = force.out.bluti_bal)


#celsius.now + celsius.12hr + month

# Computes the default residuals (raw conditional)
raw_cond_acc0 <- compute_redres(lmer_acc0_bluti_group)

#linearity - can't test as all the terms are categorical 
hist(raw_cond_acc0) #normality in histogram
plot_redres(lmer_acc0_bluti_group) #heteroskedasticity
plot_resqq(lmer_acc0_bluti_group) #normality of residuals - heavy tailed and bimodal?
plot_ranef(lmer_acc0_bluti_group) #normality of random effects residuals

simulation_output <- simulateResiduals(fittedModel = lmer_acc0_bluti_group, n = 1000)
plot(simulation_output)
```

##### robustlmm {.tab}
```{r}

lmer_acc0_bluti_group = lmer(int_acc ~ 1 + day_since*exp_group + scare_outcome + prev_scares + flight_no + age + sex + hr+ celsius.now + date_label+(1|id), data = force.out.bluti_bal)

rlmer_acc0_bluti_group = rlmer(int_acc ~ 1 + day_since*exp_group + scare_outcome + prev_scares + flight_no + age + sex + hr+ celsius.now + date_label+(1|id), data = force.out.bluti_bal)

summary(rlmer_acc0_bluti_group)
plot(rlmer_acc0_bluti_group)
```

```{r}
rlmer2_acc0_group <- update(rlmer_acc0_bluti_group, rho.sigma.e = psi2propII(smoothPsi, k = 1.66), 
                 rho.sigma.b = psi2propII(smoothPsi, k = 1.66)) #both fixed and random effects -> k = 1.66 
```

```{r}
rlmer3_acc0_group <- update(rlmer2_acc0_group,  rho.sigma.e = psi2propII(smoothPsi, k = 2.28), 
                 rho.sigma.b = psi2propII(smoothPsi, k = 2.28))
```

```{r}
compare(lmer_acc0_bluti_group, rlmer_acc0_bluti_group, rlmer2_acc0_group, rlmer3_acc0_group, show.rho.functions = FALSE)
#12.1, 13, 12.6, 12.3
```
###### tab_model
```{r}
tab_model(rlmer3_acc0_group)
```
###### plot 
```{r}
rlmer_acc0_scare_predict <- predict_response(rlmer3_acc0_group, "scare_outcome", margin = "empirical")

acc0_scare <- plot(rlmer_acc0_scare_predict) + labs(x ="", y = "acc0 (m/s^2)", title = "")
acc0_scare

rlmer_acc0_group_predict <- predict_response(rlmer3_acc0_group, c("day_since","exp_group"), margin = "empirical")
acc0_group <- plot(rlmer_acc0_group_predict)+ labs(x = "", y = "acc0 (m/s^2)", title = "", colour = "group") +scale_color_manual(values = c("#0072B2", "#CC79A7"))   
 
acc0_group

rlmer_acc0_group_interact <- predict_response(rlmer3_acc0_group, c("day_since","exp_group"), margin = "marginalmeans")
```
###### test
```{r}
emm_options(rg.limit = 40000)

#scare
test_predictions(rlmer_acc0_scare_predict)

#group

test_predictions(rlmer_acc0_group_predict, collapse_levels = TRUE, p_adjust = "sidak")
#treatment significant for day0
#control -1 vs. 0 <0.001
#control 0 vs. 1 0.072

test_predictions(rlmer_acc0_group_interact, test = "interaction", p_adjust = "sidak") 
#-1 vs. 0  0.063
# 0 vs. 1  0.043
```
#### v1
##### model assumption
```{r}

lmer_v1_bluti_group = lmer(v1 ~ 1  + day_since*exp_group + scare_outcome + prev_scares + flight_no  + sex + age + hr+ celsius.now + date_label + (1|id), data = ir.out.bluti_bal)


#celsius.now + celsius.12hr + month

# Computes the default residuals (raw conditional)
raw_cond_v1 <- compute_redres(lmer_v1_bluti_group)

#linearity - can't test as all the terms are categorical 
hist(raw_cond_v1) #normality in histogram
plot_redres(lmer_v1_bluti_group) #heteroskedasticity
plot_resqq(lmer_v1_bluti_group) #normality of residuals - heavy tailed and bimodal?
plot_ranef(lmer_v1_bluti_group) #normality of random effects residuals

simulation_output <- simulateResiduals(fittedModel = lmer_v1_bluti_group, n = 1000)
plot(simulation_output)
```

##### robustlmm {.tab}
```{r}

lmer_v1_bluti_group = lmer(v1 ~ 1 + day_since*exp_group + scare_outcome + prev_scares + flight_no + age + sex + hr+ celsius.now + date_label+(1|id), data = ir.out.bluti_bal)

rlmer_v1_bluti_group = rlmer(v1 ~ 1 + day_since*exp_group + scare_outcome + prev_scares + flight_no + age + sex + hr+ celsius.now + date_label+(1|id), data = ir.out.bluti_bal)

summary(rlmer_v1_bluti_group)
plot(rlmer_v1_bluti_group)
```

```{r}
rlmer2_v1_group <- update(rlmer_v1_bluti_group, rho.sigma.e = psi2propII(smoothPsi, k = 1.66), 
                 rho.sigma.b = psi2propII(smoothPsi, k = 1.66)) #both fixed and random effects -> k = 1.66 
```

```{r}
rlmer3_v1_group <- update(rlmer2_v1_group,  rho.sigma.e = psi2propII(smoothPsi, k = 2.28), 
                 rho.sigma.b = psi2propII(smoothPsi, k = 2.28))
```

```{r}
compare(lmer_v1_bluti_group, rlmer_v1_bluti_group, rlmer2_v1_group, rlmer3_v1_group, show.rho.functions = FALSE)
#0.365, 0.269, 0.313, 0.355
```
###### tab_model
```{r}
tab_model(rlmer_v1_bluti_group)
```
###### plot 
```{r}
#scare
rlmer_v1_scare_predict <- predict_response(rlmer3_v1_group, "scare_outcome", margin = "empirical")

v1_scare <- plot(rlmer_v1_scare_predict) + labs(title = "", x ="", y = "v1 (m/s)")
v1_scare

#group
rlmer_v1_group_predict <- predict_response(rlmer3_v1_group, c("day_since","exp_group"),margin = "empirical")

v1_group <- plot(rlmer_v1_group_predict)+ labs(x = "day since experiment", y = "v1 (m/s)", title = "", colour = "group") +scale_color_manual(values = c("#0072B2", "#CC79A7"))  
v1_group 

rlmer_v1_group_interact <- predict_response(rlmer3_v1_group, c("day_since","exp_group"),margin = "marginalmeans")
```
###### test
```{r}
emm_options(rg.limit = 40000)
#scare

test_predictions(rlmer_v1_scare_predict)

#group
test_predictions(rlmer_v1_group_predict, collapse_levels = TRUE, p_adjust = "sidak")
#no difference within group
test_predictions(rlmer_v1_group_interact, test = "interaction", p_adust = "sidak") 
#no interaction effect
```
#### v2
##### model assumption
```{r}

lmer_v2_bluti_group = lmer(v2 ~ 1  + day_since*exp_group + scare_outcome + prev_scares + flight_no  + sex + age + hr+ celsius.now + date_label + (1|id), data = ir.out.bluti_bal)


#celsius.now + celsius.12hr + month

# Computes the default residuals (raw conditional)
raw_cond_v2 <- compute_redres(lmer_v2_bluti_group)

#linearity - can't test as all the terms are categorical 
hist(raw_cond_v2) #normality in histogram
plot_redres(lmer_v2_bluti_group) #heteroskedasticity
plot_resqq(lmer_v2_bluti_group) #normality of residuals - heavy tailed and bimodal?
plot_ranef(lmer_v2_bluti_group) #normality of random effects residuals

simulation_output <- simulateResiduals(fittedModel = lmer_v2_bluti_group, n = 1000)
plot(simulation_output)
```

##### robustlmm {.tab}
```{r}

lmer_v2_bluti_group = lmer(v2 ~ 1 + day_since*exp_group + scare_outcome + prev_scares + flight_no + age + sex + hr+ celsius.now + date_label+(1|id), data = ir.out.bluti_bal)

rlmer_v2_bluti_group = rlmer(v2 ~ 1 + day_since*exp_group + scare_outcome + prev_scares + flight_no + age + sex + hr+ celsius.now + date_label+(1|id), data = ir.out.bluti_bal)

summary(rlmer_v2_bluti_group)
plot(rlmer_v2_bluti_group)
```

```{r}
rlmer2_v2_group <- update(rlmer_v2_bluti_group, rho.sigma.e = psi2propII(smoothPsi, k = 1.66), 
                 rho.sigma.b = psi2propII(smoothPsi, k = 1.66)) #both fixed and random effects -> k = 1.66 
```

```{r}
rlmer3_v2_group <- update(rlmer2_v2_group,  rho.sigma.e = psi2propII(smoothPsi, k = 2.28), 
                 rho.sigma.b = psi2propII(smoothPsi, k = 2.28))
```

```{r}
compare(lmer_v2_bluti_group, rlmer_v2_bluti_group, rlmer2_v2_group, rlmer3_v2_group, show.rho.functions = FALSE)
#0.337, 0.141, 0.165, 0.284
```
###### tab_model
```{r}
tab_model(rlmer_v2_bluti_group)
```
###### plot 
```{r}
#scare outcome
rlmer_v2_scare_predict <- predict_response(rlmer_v2_bluti_group, "scare_outcome", margin = "empirical")

v2_scare <- plot(rlmer_v2_scare_predict) + labs(title = "", x ="", y = "v2 (m/s)")
v2_scare



#group-level
rlmer_v2_group_predict <- predict_response(rlmer_v2_bluti_group, c("day_since","exp_group"),margin = "empirical")

v2_group <- plot(rlmer_v2_group_predict)+ labs(x = "day since experiment", y = "v2 (m/s)", title = "", colour = "group") +scale_color_manual(values = c("#0072B2", "#CC79A7"))  
v2_group 

rlmer_v2_group_interact <- predict_response(rlmer_v2_bluti_group, c("day_since","exp_group"),margin = "marginalmeans")
```
###### test
```{r}
emm_options(rg.limit = 40000)
#scare
test_predictions(rlmer_v2_scare_predict)


#group
test_predictions(rlmer_v2_group_predict, collapse_levels = TRUE, p_adjust = "sidak")

#no difference within group
test_predictions(rlmer_v2_group_interact, test = "interaction", p_adjust = "sidak") 
#no interaction effect
```
#### v3
##### model assumption
```{r}

lmer_v3_bluti_group = lmer(v3 ~ 1  + day_since*exp_group + scare_outcome + prev_scares + flight_no  + sex + age + hr+ celsius.now + date_label + (1|id), data = ir.out.bluti_bal)


#celsius.now + celsius.12hr + month

# Computes the default residuals (raw conditional)
raw_cond_v3 <- compute_redres(lmer_v3_bluti_group)

#linearity - can't test as all the terms are categorical 
hist(raw_cond_v3) #normality in histogram
plot_redres(lmer_v3_bluti_group) #heteroskedasticity
plot_resqq(lmer_v3_bluti_group) #normality of residuals - heavy tailed and bimodal?
plot_ranef(lmer_v3_bluti_group) #normality of random effects residuals

simulation_output <- simulateResiduals(fittedModel = lmer_v3_bluti_group, n = 1000)
plot(simulation_output)
```

##### robustlmm {.tab}
```{r}

lmer_v3_bluti_group = lmer(v3 ~ 1 + day_since*exp_group + scare_outcome + prev_scares + flight_no + age + sex + hr+ celsius.now + date_label+(1|id), data = ir.out.bluti_bal)

rlmer_v3_bluti_group = rlmer(v3 ~ 1 + day_since*exp_group + scare_outcome + prev_scares + flight_no + age + sex + hr+ celsius.now + date_label+(1|id), data = ir.out.bluti_bal)

summary(rlmer_v3_bluti_group)
#probably singular
plot(rlmer_v3_bluti_group)
```

```{r}
rlmer2_v3_group <- update(rlmer_v3_bluti_group, rho.sigma.e = psi2propII(smoothPsi, k = 1.66), 
                 rho.sigma.b = psi2propII(smoothPsi, k = 1.66)) #both fixed and random effects -> k = 1.66
#lots of warning
```

```{r}
rlmer3_v3_group <- update(rlmer2_v3_group,  rho.sigma.e = psi2propII(smoothPsi, k = 2.28), 
                 rho.sigma.b = psi2propII(smoothPsi, k = 2.28))
```

```{r}
compare(lmer_v3_bluti_group, rlmer_v3_bluti_group, rlmer2_v3_group, rlmer3_v3_group, show.rho.functions = FALSE)
#0.17, 0.101, 0.102, 0.113
```
###### tab_model
```{r}
tab_model(rlmer_v3_bluti_group)
```
###### plot 
```{r}
#scare_outcome
rlmer_v3_scare_predict <- predict_response(rlmer_v3_bluti_group, "scare_outcome", margin = "empirical")

v3_scare <- plot(rlmer_v3_scare_predict) + labs(title = "", x ="", y = "v3 (m/s)")
 
v3_scare

#group
rlmer_v3_group_predict <- predict_response(rlmer_v3_bluti_group, c("day_since","exp_group"),margin = "empirical")

v3_group <- plot(rlmer_v3_group_predict) + labs(x = "day since experiment", y = "v3 (m/s)", title = "", colour = "group") +scale_color_manual(values = c("#0072B2", "#CC79A7")) 

v3_group 

rlmer_v3_group_interact <- predict_response(rlmer_v3_bluti_group, c("day_since","exp_group"),margin = "marginalmeans")
```
###### test
```{r}
emm_options(rg.limit = 40000)

#scare
test_predictions(rlmer_v3_scare_predict)

#group
test_predictions(rlmer_v3_group_predict, collapse_levels = TRUE, p_adjust = "sidak")
#control: -1 vs. 0 -> 0.039. estimate 0.01 m /s 
#no effect 

test_predictions(rlmer_v3_group_interact, test = "interaction", p_adjust = "sidak") 
#no interaction effect
```
#### v4
##### model assumption
```{r}

lmer_v4_bluti_group = lmer(v4 ~ 1  + day_since*exp_group + scare_outcome + prev_scares + flight_no  + sex + age + hr+ celsius.now + date_label + (1|id), data = ir.out.bluti_bal)


#celsius.now + celsius.12hr + month

# Computes the default residuals (raw conditional)
raw_cond_v4 <- compute_redres(lmer_v4_bluti_group)

#linearity - can't test as all the terms are categorical 
hist(raw_cond_v4) #normality in histogram
plot_redres(lmer_v4_bluti_group) #heteroskedasticity
plot_resqq(lmer_v4_bluti_group) #normality of residuals - heavy tailed and bimodal?
plot_ranef(lmer_v4_bluti_group) #normality of random effects residuals

simulation_output <- simulateResiduals(fittedModel = lmer_v4_bluti_group, n = 1000)
plot(simulation_output)
```

##### robustlmm {.tab}
```{r}

lmer_v4_bluti_group = lmer(v4 ~ 1 + day_since*exp_group + scare_outcome + prev_scares + flight_no + age + sex + hr+ celsius.now + date_label+(1|id), data = ir.out.bluti_bal)

rlmer_v4_bluti_group = rlmer(v4 ~ 1 + day_since*exp_group + scare_outcome + prev_scares + flight_no + age + sex + hr+ celsius.now + date_label+(1|id), data = ir.out.bluti_bal)

summary(rlmer_v4_bluti_group)
#probably singular
plot(rlmer_v4_bluti_group)
```

```{r}
rlmer2_v4_group <- update(rlmer_v4_bluti_group, rho.sigma.e = psi2propII(smoothPsi, k = 1.66), 
                 rho.sigma.b = psi2propII(smoothPsi, k = 1.66)) #both fixed and random effects -> k = 1.66
#lots of warning
```

```{r}
rlmer3_v4_group <- update(rlmer2_v4_group,  rho.sigma.e = psi2propII(smoothPsi, k = 2.28), 
                 rho.sigma.b = psi2propII(smoothPsi, k = 2.28))
```

```{r}
compare(lmer_v4_bluti_group, rlmer_v4_bluti_group, rlmer2_v4_group, rlmer3_v4_group, show.rho.functions = FALSE)
#0.424, 0.114, 0.118, 0.14
```
###### tab_model
```{r}
tab_model(rlmer_v4_bluti_group)
```
###### plot 
```{r}
#scare
rlmer_v4_scare_predict <- predict_response(rlmer_v4_bluti_group, "scare_outcome", margin = "empirical")

v4_scare <- plot(rlmer_v4_scare_predict) + labs(title = "", x ="", y="v4 (m/s)")
v4_scare

#group
rlmer_v4_group_predict <- predict_response(rlmer_v4_bluti_group, c("day_since","exp_group"),margin = "empirical")

v4_group <- plot(rlmer_v4_group_predict)+ labs(x = "day since experiment", y = "v4 (m/s)", title = "", colour = "group") +scale_color_manual(values = c("#0072B2", "#CC79A7"))

v4_group 

rlmer_v4_group_interact <- predict_response(rlmer_v4_bluti_group, c("day_since","exp_group"),margin = "marginalmeans")
```
###### test
```{r}
emm_options(rg.limit = 40000)
#scare
test_predictions(rlmer_v4_scare_predict)

#group
test_predictions(rlmer_v4_group_predict, collapse_levels = TRUE, p_adjust = "sidak")
#control: -1 vs.1 -> 0.024. estimate 0.01 m /s 
#no effect 
test_predictions(rlmer_v4_group_interact, test = "interaction", p_adjust = "sidak") 
#no interaction effect
```




#### acc1
##### model assumption
```{r}

lmer_acc1_bluti_group = lmer(acc1 ~ 1  + day_since*exp_group + scare_outcome + prev_scares + flight_no  + sex + age + hr+ celsius.now + date_label + (1|id), data = ir.out.bluti_bal)


#celsius.now + celsius.12hr + month

# Computes the default residuals (raw conditional)
raw_cond_acc1 <- compute_redres(lmer_acc1_bluti_group)

#linearity - can't test as all the terms are categorical 
hist(raw_cond_acc1) #normality in histogram
plot_redres(lmer_acc1_bluti_group) #heteroskedasticity
plot_resqq(lmer_acc1_bluti_group) #normality of residuals - heavy tailed and bimodal?
plot_ranef(lmer_acc1_bluti_group) #normality of random effects residuals

simulation_output <- simulateResiduals(fittedModel = lmer_acc1_bluti_group, n = 1000)
plot(simulation_output)
```

##### robustlmm {.tab}
```{r}

lmer_acc1_bluti_group = lmer(acc1 ~ 1 + day_since*exp_group + scare_outcome + prev_scares + flight_no + age + sex + hr+ celsius.now + date_label+(1|id), data = ir.out.bluti_bal)

rlmer_acc1_bluti_group = rlmer(acc1 ~ 1 + day_since*exp_group + scare_outcome + prev_scares + flight_no + age + sex + hr+ celsius.now + date_label+(1|id), data = ir.out.bluti_bal)

summary(rlmer_acc1_bluti_group)
#probably singular
plot(rlmer_acc1_bluti_group)
```

```{r}
rlmer2_acc1_group <- update(rlmer_acc1_bluti_group, rho.sigma.e = psi2propII(smoothPsi, k = 1.66), 
                 rho.sigma.b = psi2propII(smoothPsi, k = 1.66)) #both fixed and random effects -> k = 1.66

```

```{r}
rlmer3_acc1_group <- update(rlmer2_acc1_group,  rho.sigma.e = psi2propII(smoothPsi, k = 2.28), 
                 rho.sigma.b = psi2propII(smoothPsi, k = 2.28))
```

```{r}
compare(lmer_acc1_bluti_group, rlmer_acc1_bluti_group, rlmer2_acc1_group, rlmer3_acc1_group, show.rho.functions = FALSE)
#3.62, 3.42, 3.37, 3.44
```
###### tab_model
```{r}
tab_model(rlmer2_acc1_group)
```
###### plot 
```{r}

#scare
rlmer_acc1_scare_predict <- predict_response(rlmer2_acc1_group, "scare_outcome", margin = "empirical")

acc1_scare <- plot(rlmer_acc1_scare_predict) + labs(title = "", x = "", y = "acc1 (m/s^2)")

acc1_scare

#group

rlmer_acc1_group_predict <- predict_response(rlmer2_acc1_group, c("day_since","exp_group"),margin = "empirical")

acc1_group <- plot(rlmer_acc1_group_predict)+ labs(x = "", y = "acc1 (m/s^2)", title = "", colour = "group") +scale_color_manual(values = c("#0072B2", "#CC79A7"))

acc1_group 

rlmer_acc1_group_interact <- predict_response(rlmer2_acc1_group, c("day_since","exp_group"),margin = "marginalmeans")
```
###### test
```{r}
emm_options(rg.limit = 40000)
#scare
test_predictions(rlmer_acc1_scare_predict)

#group
test_predictions(rlmer_acc1_group_predict, collapse_levels = TRUE)

#no effect - but estimates tend to be higher for control 
test_predictions(rlmer_acc1_group_interact, test = "interaction") 
#no interaction effect
```
#### acc2
##### model assumption
```{r}

lmer_acc2_bluti_group = lmer(acc2 ~ 1  + day_since*exp_group + scare_outcome + prev_scares + flight_no  + sex + age + hr+ celsius.now + date_label + (1|id), data = ir.out.bluti_bal)


#celsius.now + celsius.12hr + month

# Computes the default residuals (raw conditional)
raw_cond_acc2 <- compute_redres(lmer_acc2_bluti_group)

#linearity - can't test as all the terms are categorical 
hist(raw_cond_acc2) #normality in histogram
plot_redres(lmer_acc2_bluti_group) #heteroskedasticity
plot_resqq(lmer_acc2_bluti_group) #normality of residuals - heavy tailed and bimodal?
plot_ranef(lmer_acc2_bluti_group) #normality of random effects residuals

simulation_output <- simulateResiduals(fittedModel = lmer_acc2_bluti_group, n = 1000)
plot(simulation_output)
```

##### robustlmm {.tab}
```{r}

lmer_acc2_bluti_group = lmer(acc2 ~ 1 + day_since*exp_group + scare_outcome + prev_scares + flight_no + age + sex + hr+ celsius.now + date_label+(1|id), data = ir.out.bluti_bal)

rlmer_acc2_bluti_group = rlmer(acc2 ~ 1 + day_since*exp_group + scare_outcome + prev_scares + flight_no + age + sex + hr+ celsius.now + date_label+(1|id), data = ir.out.bluti_bal)

summary(rlmer_acc2_bluti_group)
#probably singular
plot(rlmer_acc2_bluti_group)
```

```{r}
rlmer2_acc2_group <- update(rlmer_acc2_bluti_group, rho.sigma.e = psi2propII(smoothPsi, k = 1.66), 
                 rho.sigma.b = psi2propII(smoothPsi, k = 1.66)) #both fixed and random effects -> k = 1.66

```

```{r}
rlmer3_acc2_group <- update(rlmer2_acc2_group,  rho.sigma.e = psi2propII(smoothPsi, k = 2.28), 
                 rho.sigma.b = psi2propII(smoothPsi, k = 2.28))
```

```{r}
compare(lmer_acc2_bluti_group, rlmer_acc2_bluti_group, rlmer2_acc2_group, rlmer3_acc2_group, show.rho.functions = FALSE)
#3.64, 2.88, 3.09, 3.55
```
###### tab_model
```{r}
tab_model(rlmer_acc2_bluti_group)
```
###### plot 
```{r}
#scare_outcome

rlmer_acc2_scare_predict <- predict_response(rlmer_acc2_bluti_group, "scare_outcome", margin = "empirical")

acc2_scare <- plot(rlmer_acc2_scare_predict) + labs(title = "", x = "", y = "acc2 (m/s^2)")

acc2_scare

#group
rlmer_acc2_bluti_predict <- predict_response(rlmer_acc2_bluti_group, c("day_since","exp_group"),margin = "empirical")

acc2_group <- plot(rlmer_acc2_bluti_predict)+ labs(x = "", y = "acc2 (m/s^2)", title = "", colour = "group") +scale_color_manual(values = c("#0072B2", "#CC79A7"))

acc2_group 

rlmer_acc2_bluti_interact <- predict_response(rlmer_acc2_bluti_group, c("day_since","exp_group"),margin = "marginalmeans")
```
###### test
```{r}
emm_options(rg.limit = 40000)
#scare
test_predictions(rlmer_acc2_scare_predict)


#group
test_predictions(rlmer_acc2_bluti_predict, collapse_levels = TRUE, p_adjust = "sidak")

#no effect - but estimates tend to be higher for treatment 
test_predictions(rlmer_acc2_bluti_interact, test = "interaction", p_adjust = "sidak") 
#no interaction effect
```
#### acc3
##### model assumption
```{r}

lmer_acc3_bluti_group = lmer(acc3 ~ 1  + day_since*exp_group + scare_outcome + prev_scares + flight_no  + sex + age + hr+ celsius.now + date_label + (1|id), data = ir.out.bluti_bal)


#celsius.now + celsius.12hr + month

# Computes the default residuals (raw conditional)
raw_cond_acc3 <- compute_redres(lmer_acc3_bluti_group)

#linearity - can't test as all the terms are categorical 
hist(raw_cond_acc3) #normality in histogram
plot_redres(lmer_acc3_bluti_group) #heteroskedasticity
plot_resqq(lmer_acc3_bluti_group) #normality of residuals - heavy tailed and bimodal?
plot_ranef(lmer_acc3_bluti_group) #normality of random effects residuals

simulation_output <- simulateResiduals(fittedModel = lmer_acc3_bluti_group, n = 1000)
plot(simulation_output)
```

##### robustlmm {.tab}
```{r}
#singular
lmer_acc3_bluti_group = lmer(acc3 ~ 1 + day_since*exp_group + scare_outcome + prev_scares + flight_no + age + sex + hr+ celsius.now + date_label+(1|id), data = ir.out.bluti_bal)
#probably singular

rlmer_acc3_bluti_group = rlmer(acc3 ~ 1 + day_since*exp_group + scare_outcome + prev_scares + flight_no + age + sex + hr+ celsius.now + date_label+(1|id), data = ir.out.bluti_bal)

summary(rlmer_acc3_bluti_group)
#probably singular
plot(rlmer_acc3_bluti_group)
```

```{r}
rlmer2_acc3_group <- update(rlmer_acc3_bluti_group, rho.sigma.e = psi2propII(smoothPsi, k = 1.66), 
                 rho.sigma.b = psi2propII(smoothPsi, k = 1.66)) #both fixed and random effects -> k = 1.66
#50 warnings
```

```{r}
rlmer3_acc3_group <- update(rlmer2_acc3_group,  rho.sigma.e = psi2propII(smoothPsi, k = 2.28), 
                 rho.sigma.b = psi2propII(smoothPsi, k = 2.28))
```

```{r}
compare(lmer_acc3_bluti_group, rlmer_acc3_bluti_group, rlmer2_acc3_group, rlmer3_acc3_group, show.rho.functions = FALSE)
#2.34, 1.94, 1.96, 2.13
```

###### tab_model
```{r}
tab_model(rlmer_acc3_bluti_group)
```
###### plot 
```{r}

# scare
rlmer_acc3_scare_predict <- predict_response(rlmer_acc3_bluti_group, "scare_outcome", margin = "empirical")

acc3_scare <- plot(rlmer_acc3_scare_predict) + labs(title = "", x = "", y = "acc3 (m/s^2)")

acc3_scare

#group
rlmer_acc3_group_predict <- predict_response(rlmer_acc3_bluti_group, c("day_since","exp_group"),margin = "empirical")

acc3_group <- plot(rlmer_acc3_group_predict)+ labs(x = "", y = "acc3 (m/s^2)", title = "", colour = "group") +scale_color_manual(values = c("#0072B2", "#CC79A7"))

acc3_group 

rlmer_acc3_group_interact <- predict_response(rlmer_acc3_bluti_group, c("day_since","exp_group"),margin = "marginalmeans")
```
###### test
```{r}
emm_options(rg.limit = 40000)
#scare
test_predictions(rlmer_acc3_scare_predict)

#group
test_predictions(rlmer_acc3_group_predict, collapse_levels = TRUE, p_adjust = "sidak")

#no effect - but estimates tend to be higher for treatment 

test_predictions(rlmer_acc3_group_interact, test = "interaction", p_adjust = "sidak")

#no interaction effect
```

### PLOTS
#### scare_outcome
```{r}

ggarrange(bw_scare, maxF_scare, Fx_scare, Fy_scare,
          acc0_scare, acc1_scare, acc2_scare, acc3_scare,
          v1_scare, v2_scare, v3_scare,v4_scare, nrow = 3, ncol = 4)

```
#### group comparisons
```{r}
ggarrange(bw_group, maxF_group, Fx_group, Fy_group,
          acc0_group, acc1_group, acc2_group, acc3_group,
          v1_group, v2_group, v3_group,v4_group, nrow = 3, ncol = 4)

```
### GRETI  
#### bodyweight {.tabset}  
##### model assumptions  
```{r}

lmer_bw_greti = lmer(bodyweight ~ 1 + day_since*group  +scare_outcome +prev_scares + hr + sex + age+ celsius.now  +  (1|id:date)+(1|id), data = force.out.greti_bal)

# Computes the default residuals (raw conditional)
raw_cond_bw <- compute_redres(lmer_bw_greti)

#linearity - can't test as all the terms are categorical 

hist(raw_cond_bw) #normality in histogram
plot_redres(lmer_bw_greti) #heteroskedasticity
plot_resqq(lmer_bw_greti) #normality of residuals - heavy tailed and bimodal?

plot_ranef(lmer_bw_greti) #normality of random effects residuals


```

##### model analysis  
```{r}
#Anova(lmer_bw_greti, type = 3, test.statistic = "F")
tab_model(lmer_bw_greti)
summary(lmer_bw_greti)
lsmeans_bw_greti <- lsmeans(lmer_bw_greti, pairwise~day_since*group)

#,lmerTest.limit = 7000
#,lmerTest.limit = 7000
```

#### peak_force {.tabset}
##### model assumptions
```{r}
lmer_peakF_greti = lmer(max_Ftot ~ 1 + day_since*group + scare_outcome + prev_scares + hr + sex + age + celsius.now + (1|id:date)+(1|id), data = force.out.greti_bal)

raw_cond_peakF <- compute_redres(lmer_peakF_greti)

hist(raw_cond_peakF) #normality in histogram
plot_redres(lmer_peakF_greti) #heteroskedasticity
plot_resqq(lmer_peakF_greti) #normality of residuals - heavy tailed and bimodal?

plot_ranef(lmer_peakF_greti) #normality of random effects residuals



```
##### model analysis
```{r}

summary(lmer_peakF_greti)
lsmeans_peakF_greti <- lsmeans(lmer_peakF_greti, pairwise~day_since*group)
tab_model(lmer_peakF_greti)


```

#### acc0 {.tabset}  
##### model assumptions  
```{r}
lmer_acc0_greti = lmer(int_acc ~ 1 + day_since*group  + scare_outcome + prev_scares + hr + sex + age +  celsius.now  + (1|id:date)+(1|id), data = force.out.greti_bal)

# Computes the default residuals (raw conditional)
raw_cond_acc0 <- compute_redres(lmer_acc0_greti)

#linearity - can't test as all the terms are categorical 

hist(raw_cond_acc0) #normality in histogram
plot_redres(lmer_acc0_greti) #heteroskedasticity
plot_resqq(lmer_acc0_greti) #normality of residuals - heavy tailed and bimodal?

plot_ranef(lmer_acc0_greti) #normality of random effects residuals

```
##### model analysis  
```{r}
#Anova(lmer_acc0_greti, type = 3, test.statistic = "F") 
summary(lmer_acc0_greti)
tab_model(lmer_acc0_greti)
lsmeans_acc0_greti <- lsmeans(lmer_acc0_greti, pairwise~day_since*group)

#,lmerTest.limit = 7000
#,lmerTest.limit = 7000
```

#### acc1 {.tabest}  
##### model assumptions  
```{r}
ir.out.greti_bal2 <- ir.out.greti_bal %>%
  filter(-30 < acc1 & acc1 <30) %>%
  filter(-20 < acc2 & acc2 <20) %>%
  filter(-20 < acc3 & acc3 <20) %>%
  na.omit
  #filter(day_since != "-1") %>%
  


lmer_acc1_greti = lmer(acc1 ~ 1 + day_since*group  + scare_outcome + prev_scares + hr + sex + age +  celsius.now + (1|id:date)+(1|id), data = ir.out.greti_bal2)

# Computes the default residuals (raw conditional)
raw_cond_acc1 <- compute_redres(lmer_acc1_greti)

#linearity - can't test as all the terms are categorical 

hist(raw_cond_acc1) #normality in histogram
plot_redres(lmer_acc1_greti) #heteroskedasticity
plot_resqq(lmer_acc1_greti) #normality of residuals - heavy tailed and bimodal?

plot_ranef(lmer_acc1_greti) #normality of random effects residuals

```
##### model analysis  
```{r}

#Anova(lmer_acc1_greti, type = 3, test.statistic = "F") 
tab_model(lmer_acc1_greti)
summary(lmer_acc1_greti)
lsmeans_acc1_greti <- lsmeans(lmer_acc1_greti, pairwise~day_since*group)

#,lmerTest.limit = 7000
#,lmerTest.limit = 7000
```

#### acc2 {.tabset}  
##### model assumptions  
```{r}
lmer_acc2_greti = lmer(acc2 ~ 1 + day_since*group  + scare_outcome + prev_scares + hr + sex + age +  celsius.now+ (1|id:date)+(1|id), data = ir.out.greti_bal2)

# Computes the default residuals (raw conditional)
raw_cond_acc2 <- compute_redres(lmer_acc2_greti)

#linearity - can't test as all the terms are categorical 

hist(raw_cond_acc2) #normality in histogram
plot_redres(lmer_acc2_greti) #heteroskedasticity
plot_resqq(lmer_acc2_greti) #normality of residuals - heavy tailed and bimodal?

plot_ranef(lmer_acc2_greti) #normality of random effects residuals

```
##### model analysis  
```{r}
#Anova(lmer_acc2_greti, type = 3, test.statistic = "F") 
tab_model(lmer_acc2_greti)
summary(lmer_acc2_greti)
lsmeans_acc2_greti <- lsmeans(lmer_acc2_greti, pairwise~day_since*group)

#,lmerTest.limit = 7000
#,lmerTest.limit = 7000

```


#### acc3 {.tabset}  
##### model assumptions  
```{r}
lmer_acc3_greti = lmer(acc3 ~ 1 + day_since*group  + scare_outcome + prev_scares + hr + sex + age +  celsius.now + (1|date), data = ir.out.greti_bal2)
#(1/id) or (1/id:date) all singular
# Computes the default residuals (raw conditional)
raw_cond_acc3 <- compute_redres(lmer_acc3_greti)

#linearity - can't test as all the terms are categorical 

hist(raw_cond_acc3) #normality in histogram
plot_redres(lmer_acc3_greti) #heteroskedasticity
plot_resqq(lmer_acc3_greti) #normality of residuals - heavy tailed and bimodal?

plot_ranef(lmer_acc3_greti) #normality of random effects residuals

```

##### model analysis  
```{r}
#Anova(lmer_acc3_greti, type = 3, test.statistic = "F") 
tab_model(lmer_acc3_greti)
summary(lmer_acc3_greti)
lsmeans_acc3_greti <- lsmeans(lmer_acc3_greti, pairwise~day_since*group)

#,lmerTest.limit = 7000
#,lmerTest.limit = 7000

```
## D. Cluster-robust variance estimaters
### BLUTI
#### bodyweight
```{r}
plm_bw_bluti <- plm(bodyweight ~ 0 + day_since*group + scare_outcome + prev_scares + hr + sex + age + celsius.now + flight_no, data = force.out.bluti_bal, effect = "individual", index = c("id", "date"), model = "random")

coef_test(plm_bw_bluti, vcov = "CR2", test = "Satterthwaite")
```


#### acc0
```{r}
plm_acc0_bluti <- plm(int_acc ~ 0 + day_since*group + scare_outcome + prev_scares + hr + sex + age + celsius.now + flight_no, data = force.out.bluti_bal, effect = "individual", index = c("id", "date"), model = "random")

coef_test(plm_acc0_bluti, vcov = "CR2", test = "Satterthwaite")

```

## E. graphs  
### plot data  
```{r}
plot.bw.bl <- data.frame(lsmeans_bw_bluti$lsmeans)
plot.peakF.bl_scare <- data.frame(lsmeans_peakF_bluti_scare$lsmeans)
plot.peakF.bl_daygroup <- data.frame(lsmeans_peakF_bluti_daygroup$lsmeans)

plot.Fx.bl_scare <- data.frame(lsmeans_Fx_bluti_scare$lsmeans)
plot.Fx.bl_daygroup <- data.frame(lsmeans_Fx_bluti_daygroup$lsmeans)

plot.Fy.bl_scare <- data.frame(lsmeans_Fy_bluti_scare$lsmeans)
plot.Fy.bl_daygroup <- data.frame(lsmeans_Fy_bluti_daygroup$lsmeans)

plot.acc0.bl <- data.frame(lsmeans_acc0_bluti$lsmeans)
plot.acc1.bl <- data.frame(lsmeans_acc1_bluti$lsmeans)
plot.acc2.bl <- data.frame(lsmeans_acc2_bluti$lsmeans)
plot.acc3.bl <- data.frame(lsmeans_acc3_bluti$lsmeans)

plot.bw.gr <- data.frame(lsmeans_bw_greti$lsmeans)
plot.peakF.gr <- data.frame(lsmeans_peakF_greti$lsmeans)
plot.acc0.gr <- data.frame(lsmeans_acc0_greti$lsmeans)
plot.acc1.gr <- data.frame(lsmeans_acc1_greti$lsmeans)
plot.acc2.gr <- data.frame(lsmeans_acc2_greti$lsmeans)
plot.acc3.gr <- data.frame(lsmeans_acc3_greti$lsmeans)
```


### graphs    
#### BLUTI {.tabset}  
##### save plots   
```{R}
plot.lsmeans_bw.bl <- 
  ggplot() +  
  geom_point(data = emmeans_bw_bluti_df, aes(x = day_since, y = emmean, alpha = 1)) +
  geom_errorbar(data = plot.bw.bl, aes(x = day_since, ymin = lower.CL, ymax = upper.CL, width = 0.5),alpha = 1) +
  ylab("bodyweight (g)-bluti") +
  #ylim(0.014,0.018) +
  theme(axis.title.x=element_blank()) + #getting rid of x-axis title
  theme(legend.position = "none") +
  facet_grid(~group)

plot.lsmeans_peakF.bl_daygroup <- 
  ggplot() +  
  geom_point(data = plot.peakF.bl_daygroup, aes(x = day_since, y = lsmean, alpha = 1)) +
  geom_errorbar(data = plot.peakF.bl_daygroup, aes(x = day_since, ymin = asymp.LCL, ymax = asymp.UCL, width = 0.5),alpha = 1) +
  ylab("peak force (N)") +
  #ylim(0.014,0.018) +
  theme(axis.title.x=element_blank()) + #getting rid of x-axis title
  theme(legend.position = "none") +
  facet_grid(~group)

plot.lsmeans_peakF.bl_scare <- 
  ggplot() +  
  geom_point(data = plot.peakF.bl_scare, aes(x = as.factor(scare_outcome), y = lsmean, alpha = 1)) +
  geom_errorbar(data = plot.peakF.bl_scare, aes(x = scare_outcome, ymin = asymp.LCL, ymax = asymp.UCL, width = 0.5),alpha = 1) +
  ylab("peak force (N)") +
  #ylim(0.014,0.018) +
  scale_x_discrete(labels = c("normal", "escape"))+
  theme(axis.title.x=element_blank()) + #getting rid of x-axis title
  theme(legend.position = "none") 

plot.lsmeans_Fx.bl_daygroup <- 
  ggplot() +  
  geom_point(data = plot.Fx.bl_daygroup, aes(x = day_since, y = lsmean, alpha = 1)) +
  geom_errorbar(data = plot.Fx.bl_daygroup, aes(x = day_since, ymin = asymp.LCL, ymax = asymp.UCL, width = 0.5),alpha = 1) +
  ylab("vertical force (N)") +
  #ylim(0.014,0.018) +
  theme(axis.title.x=element_blank()) + #getting rid of x-axis title
  theme(legend.position = "none") +
  facet_grid(~group)

plot.lsmeans_Fx.bl_scare <- 
  ggplot() +  
  geom_point(data = plot.Fx.bl_scare, aes(x = as.factor(scare_outcome), y = lsmean, alpha = 1)) +
  geom_errorbar(data = plot.Fx.bl_scare, aes(x = scare_outcome, ymin = asymp.LCL, ymax = asymp.UCL, width = 0.5),alpha = 1) +
  ylab("vertical force (N)") +
  #ylim(0.014,0.018) +
  scale_x_discrete(labels = c("normal", "escape"))+
  theme(axis.title.x=element_blank()) + #getting rid of x-axis title
  theme(legend.position = "none") 


plot.lsmeans_Fy.bl_daygroup <- 
  ggplot() +  
  geom_point(data = plot.Fy.bl_daygroup, aes(x = day_since, y = -lsmean, alpha = 1)) +
  geom_errorbar(data = plot.Fy.bl_daygroup, aes(x = day_since, ymin = -asymp.LCL, ymax = -asymp.UCL, width = 0.5),alpha = 1) +
  ylab("horizontal force (N)") +
  #ylim(0.014,0.018) +
  theme(axis.title.x=element_blank()) + #getting rid of x-axis title
  theme(legend.position = "none") +
  facet_grid(~group)

plot.lsmeans_Fy.bl_scare <- 
  ggplot() +  
  geom_point(data = plot.Fy.bl_scare, aes(x = as.factor(scare_outcome), y = -lsmean, alpha = 1)) +
  geom_errorbar(data = plot.Fy.bl_scare, aes(x = scare_outcome, ymin = -asymp.LCL, ymax = -asymp.UCL, width = 0.5),alpha = 1) +
  ylab("horizontal force (N)") +
  #ylim(0.014,0.018) +
  scale_x_discrete(labels = c("normal", "takeoff"))+
  theme(axis.title.x=element_blank()) + #getting rid of x-axis title
  theme(legend.position = "none") 



plot.lsmeans0.bl <- 
  ggplot() +  
  geom_point(data = plot.acc0.bl, aes(x = day_since, y = lsmean, alpha = 1)) +
  geom_errorbar(data = plot.acc0.bl, aes(x = day_since, ymin = asymp.LCL, ymax = asymp.UCL, width = 0.5),alpha = 1) +
    ylab(expression(inital~leg~acceleration~(m/s^2)-bluti)) +
  theme(axis.title.x=element_blank()) + #getting rid of x-axis title
  theme(legend.position = "none") +
  facet_grid(~group)


plot.lsmeans1.bl <- 
  ggplot() +  
  geom_point(data = plot.acc1.bl, aes(x = day_since, y = lsmean), alpha = 1) +
  geom_errorbar(data = plot.acc1.bl, aes(x = day_since, ymin = asymp.LCL, ymax = asymp.UCL, width = 0.5),alpha = 1) +
  ylab(expression(acc~1~(m/s^2)-bluti)) +
  #ylim(-2,6) +
  theme(axis.title.x=element_blank()) + #getting rid of x-axis title
  theme(legend.position = "none") +
  facet_grid(~group)



plot.lsmeans2.bl <- 
  ggplot() +  
  geom_point(data = plot.acc2.bl, aes(x = day_since, y = lsmean), alpha = 1) +
  geom_errorbar(data = plot.acc2.bl, aes(x = day_since, ymin = asymp.LCL, ymax = asymp.UCL, width = 0.5),alpha = 1) +
  ylab(expression(acc~2~(m/s^2)-bluti)) +
  #ylim(-2,6) +
  theme(axis.title.x=element_blank()) + #getting rid of x-axis title
  theme(legend.position = "none") +
  facet_grid(~group)


plot.lsmeans3.bl <- 
  ggplot() +  
  geom_point(data = plot.acc3.bl, aes(x = day_since, y = lsmean), alpha = 1) +
  geom_errorbar(data = plot.acc3.bl, aes(x = day_since, ymin = asymp.LCL, ymax = asymp.UCL, width = 0.5),alpha = 1) +
  ylab(expression(acc~3~(m/s^2)-bluti)) +
  #ylim(-2,6) +
  theme(axis.title.x=element_blank()) + #getting rid of x-axis title
  theme(legend.position = "none") +
  facet_grid(~group)
```
##### show plots  
```{r}

grid.arrange(plot.lsmeans_peakF.bl_daygroup, plot.lsmeans_peakF.bl_scare, plot.lsmeans_Fx.bl_daygroup,  plot.lsmeans_Fx.bl_scare, plot.lsmeans_Fy.bl_daygroup, plot.lsmeans_Fy.bl_scare, nrow = 3, top = "Model blue tit leg take-off force")
grid.arrange(plot.lsmeans_bw.bl, plot.lsmeans0.bl, plot.lsmeans1.bl, plot.lsmeans2.bl, plot.lsmeans3.bl,nrow = 3)

```

#### GRETI {.tabset}  
##### save plots  
```{r}
plot.lsmeans_bw.gr <- 
  ggplot() +  
  geom_point(data = plot.bw.gr, aes(x = day_since, y = lsmean, alpha = 1)) +
  geom_errorbar(data = plot.bw.gr, aes(x = day_since, ymin = lower.CL, ymax = upper.CL, width = 0.5),alpha = 1) +
  ylab("bodyweight (g)-greti") +
  #ylim(0.014,0.018) +
  theme(axis.title.x=element_blank()) + #getting rid of x-axis title
  theme(legend.position = "none") +
  facet_grid(~group)

plot.lsmeans_peakF.gr <- 
  ggplot() +  
  geom_point(data = plot.peakF.gr, aes(x = day_since, y = lsmean, alpha = 1)) +
  geom_errorbar(data = plot.peakF.gr, aes(x = day_since, ymin = lower.CL, ymax = upper.CL, width = 0.5),alpha = 1) +
  ylab("peak force (N)-greti") +
  #ylim(0.014,0.018) +
  theme(axis.title.x=element_blank()) + #getting rid of x-axis title
  theme(legend.position = "none") +
  facet_grid(~group)
  



plot.lsmeans0.gr <- 
  ggplot() +  
  geom_point(data = plot.acc0.gr, aes(x = day_since, y = lsmean, alpha = 1)) +
  geom_errorbar(data = plot.acc0.gr, aes(x = day_since, ymin = lower.CL, ymax = upper.CL, width = 0.5),alpha = 1) +
    ylab(expression(inital~leg~acceleration~(m/s^2)-greti)) +
  theme(axis.title.x=element_blank()) + #getting rid of x-axis title
  theme(legend.position = "none") +
  facet_grid(~group)



plot.lsmeans1.gr <- 
  ggplot() +  
  geom_point(data = plot.acc1.gr, aes(x = day_since, y = lsmean), alpha = 1) +
  geom_errorbar(data = plot.acc1.gr, aes(x = day_since, ymin = lower.CL, ymax = upper.CL, width = 0.5),alpha = 1) +
  ylab(expression(acc~1~(m/s^2)-greti)) +
  #ylim(-2,6) +
  theme(axis.title.x=element_blank()) + #getting rid of x-axis title
  theme(legend.position = "none") +
  facet_grid(~group)


plot.lsmeans2.gr <- 
  ggplot() +  
  geom_point(data = plot.acc2.gr, aes(x = day_since, y = lsmean), alpha = 1) +
  geom_errorbar(data = plot.acc2.gr, aes(x = day_since, ymin = lower.CL, ymax = upper.CL, width = 0.5),alpha = 1) +
  ylab(expression(acc~2~(m/s^2)-greti)) +
  #ylim(-2,6) +
  theme(axis.title.x=element_blank()) + #getting rid of x-axis title
  theme(legend.position = "none") +
  facet_grid(~group)



plot.lsmeans3.gr <- 
  ggplot() +  
  geom_point(data = plot.acc3.gr, aes(x = day_since, y = lsmean), alpha = 1) +
  geom_errorbar(data = plot.acc3.gr, aes(x = day_since, ymin = lower.CL, ymax = upper.CL, width = 0.5),alpha = 1) +
  ylab(expression(acc~3~(m/s^2)-greti)) +
  #ylim(-2,6) +
  theme(axis.title.x=element_blank()) + #getting rid of x-axis title
  theme(legend.position = "none") +
  facet_grid(~group)


```
##### show plots  
```{r}
grid.arrange(plot.lsmeans_bw.gr, plot.lsmeans_peakF.gr,plot.lsmeans0.gr, plot.lsmeans1.gr, plot.lsmeans2.gr, plot.lsmeans3.gr, nrow = 2)
```



## E. Best models selection {.tabset}  
Best models are not running because maybe too many terms.. 
### BLUTI {.tabset}  
#### bodyweight  
```{r, warning = FALSE}
#force.out.bluti.complete <- force.out.bluti_bal2[complete.cases(force.out.bluti_bal2), ]
summary(model.avg(dredge(lmer(bodyweight ~ 1 + day_since*group  +scare_outcome +prev_scares + hr + sex + age+ celsius.now + (1|id:date)+(1|id), data = force.out.bluti_bal, na.action = "na.fail")), subset = delta <  10))

```
#### peakF
```{r}
full_model <- lmer(max_Ftot ~ 1 + day_since*exp_group  +scare_outcome +prev_scares + flight_no + sex + age +hr +  celsius.now +date_label + (1|id), data = force.out.bluti_bal, na.action = "na.fail")

check_collinearity(full_model)
#date turned out to be correlated with day_since so dropping date into random effects

summary(model.avg(dredge(lmer(max_Ftot ~ 1 + day_since*exp_group  +scare_outcome +prev_scares + flight_no + sex + age +hr + celsius.now +date_label + (1|id), data = force.out.bluti_bal, na.action = "na.fail")), subset = delta <  10))




```

#### int_acc (acc0)  
```{r, message = FALSE,warning = FALSE}
summary(model.avg(dredge(lmer(int_acc ~ 1 + day_since*group  +scare_outcome +prev_scares + hr + sex + age+ celsius.now + (1|id:date)+(1|id), data = force.out.bluti_bal, na.action = "na.fail")), subset = delta <  10))

```
#### acc1  
```{r, message = FALSE, warning = FALSE}
summary(model.avg(dredge(lmer(acc1 ~ 1 + day_since*group  +scare_outcome +prev_scares + hr + sex + age+ celsius.now + (1|id:date)+(1|id), data = ir.out.bluti_bal2, na.action = "na.fail")), subset = delta <  10))
```
#### acc2  
```{r, message = FALSE, warning = FALSE}
summary(model.avg(dredge(lmer(acc2 ~ 1 + day_since*group  +scare_outcome +prev_scares + hr + sex + celsius.now + (1|id:date)+(1|id), data = ir.out.bluti_bal2, na.action = "na.fail")), subset = delta <  10))
```
#### acc3  
```{r, message = FALSE, warning = FALSE}
summary(model.avg(dredge(lmer(acc3 ~ 1 + day_since*group  +scare_outcome +prev_scares + hr + sex + age + celsius.now + (1|id:date)+(1|id), data = ir.out.bluti_bal2, na.action = "na.fail")), subset = delta <  10))
```
### GRETI {.tabset}
#### bodyweight  
```{r, warning = FALSE}

summary(model.avg(dredge(lmer(bodyweight ~ 1 + day_since*group  +scare_outcome +prev_scares + hr + sex + age+ celsius.now + (1|id:date)+(1|id), data = force.out.greti_bal, na.action = "na.fail")), subset = delta <  10))


```
#### int_acc (acc0)  
```{r, message = FALSE,warning = FALSE}
summary(model.avg(dredge(lmer(int_acc ~ 1 + day_since*group  +scare_outcome +prev_scares + hr + sex + age+ celsius.now + (1|id:date)+(1|id), data = force.out.greti_bal, na.action = "na.fail")), subset = delta <  10))

```
#### acc1  
```{r, message = FALSE, warning = FALSE}
summary(model.avg(dredge(lmer(acc1 ~ 1 + day_since*group  +scare_outcome +prev_scares + hr + sex + age+ celsius.now + (1|id:date)+(1|id), data = ir.out.greti_bal2, na.action = "na.fail")), subset = delta <  10))
```
#### acc2  
```{r, message = FALSE, warning = FALSE}
summary(model.avg(dredge(lmer(acc2 ~ 1 + day_since*group  +scare_outcome +prev_scares + hr + sex + age+ celsius.now + (1|id:date)+(1|id), data = ir.out.greti_bal2, na.action = "na.fail")), subset = delta <  10))
```
#### acc3  
```{r, message = FALSE, warning = FALSE}
summary(model.avg(dredge(lmer(acc3 ~ 1 + day_since*group  +scare_outcome +prev_scares + hr + sex + age+ celsius.now  + (1|date), data = ir.out.greti_bal2, na.action = "na.fail")), subset = delta <  10))
```

## F. Best model analysis
### BLUTI {.tabset}
#### bodyweight {.tabset}
##### model assumptions  
```{r}

lmer_bw_bluti = lmer(bodyweight ~ 1 + age + hr + scare_outcome + (1|id:date)+(1|id), data = force.out.bluti_bal)

# Computes the default residuals (raw conditional)
raw_cond_bw <- compute_redres(lmer_bw_bluti)

#linearity - can't test as all the terms are categorical 

hist(raw_cond_bw) #normality in histogram
plot_redres(lmer_bw_bluti) #heteroskedasticity
plot_resqq(lmer_bw_bluti) #normality of residuals - heavy tailed and bimodal?

plot_ranef(lmer_bw_bluti) #normality of random effects residuals


```
##### model analysis  
```{r}
#Anova(lmer_bw_bluti, type = 3, test.statistic = "F") 
#emm_options(pbkrtest.limit = 7000)
tab_model(lmer_bw_bluti)
lsmeans_bw_bluti <- lsmeans(lmer_bw_bluti, pairwise~scare_outcome)

#,lmerTest.limit = 7000
#,lmerTest.limit = 7000
```


#### acc0 {.tabset}
##### model assumptions  
```{r}

lmer_acc0_bluti = lmer(int_acc ~ 1 + day_since*group + hr+ age + sex + scare_outcome+ (1|id:date)+(1|id), data = force.out.bluti_bal)

# Computes the default residuals (raw conditional)
raw_cond_acc0 <- compute_redres(lmer_acc0_bluti)

#linearity - can't test as all the terms are categorical 

hist(raw_cond_acc0) #normality in histogram
plot_redres(lmer_acc0_bluti) #heteroskedasticity
plot_resqq(lmer_acc0_bluti) #normality of residuals - heavy tailed and bimodal?

plot_ranef(lmer_acc0_bluti) #normality of random effects residuals

```

##### model analysis  
```{r}
#Anova(lmer_acc0_bluti, type = 3, test.statistic = "F") 
summary(lmer_acc0_bluti)
lsmeans_acc0_bluti <- lsmeans(lmer_acc0_bluti, pairwise~day_since*group)
tab_model(lmer_acc0_bluti)

#,lmerTest.limit = 7000
#,lmerTest.limit = 7000
```

#### acc1 {.tabset}
##### model assumptions  
```{r}

lmer_acc1_bluti = lmer(acc1 ~ 1 + scare_outcome + sex+ (1|id:date)+(1|id), data = ir.out.bluti_bal2)

# Computes the default residuals (raw conditional)
raw_cond_acc1 <- compute_redres(lmer_acc1_bluti)

#linearity - can't test as all the terms are categorical 

hist(raw_cond_acc1) #normality in histogram
plot_redres(lmer_acc1_bluti) #heteroskedasticity
plot_resqq(lmer_acc1_bluti) #normality of residuals - heavy tailed and bimodal?

plot_ranef(lmer_acc1_bluti) #normality of random effects residuals

```
##### model analysis  
```{r}

#Anova(lmer_acc1_bluti, type = 3, test.statistic = "F") 
tab_model(lmer_acc1_bluti)
summary(lmer_acc1_bluti)
lsmeans_acc1_bluti <- lsmeans(lmer_acc1_bluti, pairwise~scare_outcome)

#,lmerTest.limit = 7000
#,lmerTest.limit = 7000
```

#### acc2 {.tabset}
##### model assumptions  
```{r}
lmer_acc2_bluti = lmer(acc2 ~ 1 + scare_outcome + (1|id:date)+(1|id), data = ir.out.bluti_bal2)

# Computes the default residuals (raw conditional)
raw_cond_acc2 <- compute_redres(lmer_acc2_bluti)

#linearity - can't test as all the terms are categorical 

hist(raw_cond_acc2) #normality in histogram
plot_redres(lmer_acc2_bluti) #heteroskedasticity
plot_resqq(lmer_acc2_bluti) #normality of residuals - heavy tailed and bimodal?

plot_ranef(lmer_acc2_bluti) #normality of random effects residuals

```

##### model analysis  
```{r}
#Anova(lmer_acc2_bluti, type = 3, test.statistic = "F") 
tab_model(lmer_acc2_bluti)
summary(lmer_acc2_bluti)
lsmeans_acc2_bluti <- lsmeans(lmer_acc2_bluti, pairwise~scare_outcome)

#,lmerTest.limit = 7000
#,lmerTest.limit = 7000

```


#### acc3 {.tabset}
##### model assumptions  
```{r}
lmer_acc3_bluti = lmer(acc3 ~ 1 + scare_outcome +(1|id), data = ir.out.bluti_bal2)
#(1|date) singular!

# Computes the default residuals (raw conditional)
raw_cond_acc3 <- compute_redres(lmer_acc3_bluti)

#linearity - can't test as all the terms are categorical 

hist(raw_cond_acc3) #normality in histogram
plot_redres(lmer_acc3_bluti) #heteroskedasticity
plot_resqq(lmer_acc3_bluti) #normality of residuals - heavy tailed and bimodal?

plot_ranef(lmer_acc3_bluti) #normality of random effects residuals

```

##### model analysis  
```{r}
#Anova(lmer_acc3_bluti, type = 3, test.statistic = "F") 
tab_model(lmer_acc3_bluti)
summary(lmer_acc3_bluti)
lsmeans_acc3_bluti <- lsmeans(lmer_acc3_bluti, pairwise~scare_outcome)

#,lmerTest.limit = 7000
#,lmerTest.limit = 7000

```

### GRETI  
#### bodyweight {.tabset}
##### model assumptions  
```{r}


lmer_bw_greti = lmer(bodyweight ~ 1 + age + hr + sex + (1|id:date)+(1|id), data = force.out.greti_bal)

# Computes the default residuals (raw conditional)
raw_cond_bw <- compute_redres(lmer_bw_greti)

#linearity - can't test as all the terms are categorical 

hist(raw_cond_bw) #normality in histogram
plot_redres(lmer_bw_greti) #heteroskedasticity
plot_resqq(lmer_bw_greti) #normality of residuals - heavy tailed and bimodal?

plot_ranef(lmer_bw_greti) #normality of random effects residuals


```

##### model analysis  
```{r}
#Anova(lmer_bw_greti, type = 3, test.statistic = "F")
tab_model(lmer_bw_greti)
summary(lmer_bw_greti)
#lsmeans_bw_greti <- lsmeans(lmer_bw_greti, pairwise~day_since*group)

#,lmerTest.limit = 7000
#,lmerTest.limit = 7000
```


#### acc0 {.tabset}
##### model assumptions  
```{r}
lmer_acc0_greti = lmer(int_acc ~ 1 + day_since*group  + scare_outcome + age + sex  + prev_scares + (1|id:date)+(1|id), data = force.out.greti_bal)

# Computes the default residuals (raw conditional)
raw_cond_acc0 <- compute_redres(lmer_acc0_greti)

#linearity - can't test as all the terms are categorical 

hist(raw_cond_acc0) #normality in histogram
plot_redres(lmer_acc0_greti) #heteroskedasticity
plot_resqq(lmer_acc0_greti) #normality of residuals - heavy tailed and bimodal?

plot_ranef(lmer_acc0_greti) #normality of random effects residuals

```

##### model analysis  
```{r}
#Anova(lmer_acc0_greti, type = 3, test.statistic = "F") 
summary(lmer_acc0_greti)
tab_model(lmer_acc0_greti)
lsmeans_acc0_greti <- lsmeans(lmer_acc0_greti, pairwise~age)

#,lmerTest.limit = 7000
#,lmerTest.limit = 7000
```

#### acc1 {.tabest}
##### model assumptions  
```{r}

lmer_acc1_greti = lmer(acc1 ~ 1  + (1|id:date)+(1|id), data = ir.out.greti_bal2)

# Computes the default residuals (raw conditional)
raw_cond_acc1 <- compute_redres(lmer_acc1_greti)

#linearity - can't test as all the terms are categorical 

hist(raw_cond_acc1) #normality in histogram
plot_redres(lmer_acc1_greti) #heteroskedasticity
plot_resqq(lmer_acc1_greti) #normality of residuals - heavy tailed and bimodal?

plot_ranef(lmer_acc1_greti) #normality of random effects residuals

```

##### model analysis  
```{r}

#Anova(lmer_acc1_greti, type = 3, test.statistic = "F") 
tab_model(lmer_acc1_greti)
summary(lmer_acc1_greti)
#lsmeans_acc1_greti <- lsmeans(lmer_acc1_greti, pairwise~age)

#,lmerTest.limit = 7000
#,lmerTest.limit = 7000
```

#### acc2 {.tabset}
##### model assumptions  
```{r}
lmer_acc2_greti = lmer(acc2 ~ 1 + (1|id:date)+(1|id), data = ir.out.greti_bal2)

# Computes the default residuals (raw conditional)
raw_cond_acc2 <- compute_redres(lmer_acc2_greti)

#linearity - can't test as all the terms are categorical 

hist(raw_cond_acc2) #normality in histogram
plot_redres(lmer_acc2_greti) #heteroskedasticity
plot_resqq(lmer_acc2_greti) #normality of residuals - heavy tailed and bimodal?

plot_ranef(lmer_acc2_greti) #normality of random effects residuals

```

##### model analysis  
```{r}
#Anova(lmer_acc2_greti, type = 3, test.statistic = "F") 
tab_model(lmer_acc2_greti)
summary(lmer_acc2_greti)
#lsmeans_acc2_greti <- lsmeans(lmer_acc2_greti, pairwise~age)

#,lmerTest.limit = 7000
#,lmerTest.limit = 7000

```


#### acc3 {.tabset}
##### model assumptions  
```{r}
lmer_acc3_greti = lmer(acc3 ~ 1 + age +  (1|id:date) + (1|id), data = ir.out.greti_bal2)
#(1|date) singularity
# Computes the default residuals (raw conditional)
raw_cond_acc3 <- compute_redres(lmer_acc3_greti)

#linearity - can't test as all the terms are categorical 

hist(raw_cond_acc3) #normality in histogram
plot_redres(lmer_acc3_greti) #heteroskedasticity
plot_resqq(lmer_acc3_greti) #normality of residuals - heavy tailed and bimodal?

plot_ranef(lmer_acc3_greti) #normality of random effects residuals

```

##### model analysis  
```{r}
#Anova(lmer_acc3_greti, type = 3, test.statistic = "F") 
tab_model(lmer_acc3_greti)
summary(lmer_acc3_greti)
lsmeans_acc3_greti <- lsmeans(lmer_acc3_greti, pairwise~age)

#,lmerTest.limit = 7000
#,lmerTest.limit = 7000

```

## G. graphs  

### plot data  
```{r}
effects_hr_bluti <- effects::effect(term= "hr", mod= lmer_bw_bluti)
summary(effects_hr_bluti)
bluti_hr <- as.data.frame(effects_hr_bluti)
bluti_hr

effects_hr_greti <- effects::effect(term= "hr", mod= lmer_bw_greti)
summary(effects_hr_greti)
greti_hr <- as.data.frame(effects_hr_greti)
greti_hr

plot.acc0.bl <- data.frame(lsmeans_acc0_bluti$lsmeans)
plot.acc1.bl <- data.frame(lsmeans_acc1_bluti$lsmeans)
plot.acc2.bl <- data.frame(lsmeans_acc2_bluti$lsmeans)
plot.acc3.bl <- data.frame(lsmeans_acc3_bluti$lsmeans)

#plot.bw.gr <- data.frame(lsmeans_bw_greti$lsmeans)
plot.acc0.gr <- data.frame(lsmeans_acc0_greti$lsmeans)
#plot.acc1.gr <- data.frame(lsmeans_acc1_greti$lsmeans)
#plot.acc2.gr <- data.frame(lsmeans_acc2_greti$lsmeans)
plot.acc3.gr <- data.frame(lsmeans_acc3_greti$lsmeans)
```


### graphs  
#### BLUTI  {.tabset}
##### save plots  
```{R}
plot.lsmeans_bw.bl <- ggplot() + 
  #2
geom_point(data=force.out.bluti_bal, aes(hr, bodyweight)) + 
  #3
  geom_point(data=bluti_hr, aes(x=hr, y=fit), color="grey", alpha = 1/100) +
  #4
  geom_line(data=bluti_hr, aes(x=hr, y=fit), color="blue") +
  #5
  geom_ribbon(data= bluti_hr, aes(x=hr, ymin=lower, ymax=upper), alpha= 0.3, fill="blue")+
  #6
  xlab("hours across the day")+
  ylab("bodyweight (g) - bluti")

plot.lsmeans0.bl <- 
  ggplot() +  
  geom_point(data = plot.acc0.bl, aes(x = day_since, y = lsmean, alpha = 1)) +
  geom_errorbar(data = plot.acc0.bl, aes(x = day_since, ymin = asymp.LCL, ymax = asymp.UCL, width = 0.5),alpha = 1) +
    ylab(expression(inital~leg~acceleration~(m/s^2)-bluti)) +
  theme(axis.title.x=element_blank()) + #getting rid of x-axis title
  theme(legend.position = "none") +
  facet_grid(~group)


plot.lsmeans1.bl <- 
  ggplot() +  
  geom_point(data = plot.acc1.bl, aes(x = scare_outcome, y = lsmean), alpha = 1) +
  geom_errorbar(data = plot.acc1.bl, aes(x = scare_outcome, ymin = asymp.LCL, ymax = asymp.UCL, width = 0.5),alpha = 1) +
  ylab(expression(acc~1~(m/s^2)-bluti)) +
  #ylim(-2,6) +
  theme(axis.title.x=element_blank()) + #getting rid of x-axis title
  theme(legend.position = "none") #+
  #facet_grid(~group)



plot.lsmeans2.bl <- 
  ggplot() +  
  geom_point(data = plot.acc2.bl, aes(x = scare_outcome, y = lsmean), alpha = 1) +
  geom_errorbar(data = plot.acc2.bl, aes(x = scare_outcome, ymin = asymp.LCL, ymax = asymp.UCL, width = 0.5),alpha = 1) +
  ylab(expression(acc~2~(m/s^2)-bluti)) +
  #ylim(-2,6) +
  theme(axis.title.x=element_blank()) + #getting rid of x-axis title
  theme(legend.position = "none") #+
  #facet_grid(~group)


plot.lsmeans3.bl <- 
  ggplot() +  
  geom_point(data = plot.acc3.bl, aes(x = scare_outcome, y = lsmean), alpha = 1) +
  geom_errorbar(data = plot.acc3.bl, aes(x = scare_outcome, ymin = asymp.LCL, ymax = asymp.UCL, width = 0.5),alpha = 1) +
  ylab(expression(acc~3~(m/s^2)-bluti)) +
  #ylim(-2,6) +
  theme(axis.title.x=element_blank()) + #getting rid of x-axis title
  theme(legend.position = "none") #+
  #facet_grid(~group)
```
##### show plots  
```{r}
plot.lsmeans_bw.bl

grid.arrange(plot.lsmeans0.bl, plot.lsmeans1.bl, plot.lsmeans2.bl, plot.lsmeans3.bl, nrow = 2)
```

#### GRETI {.tabset}
##### save plots  
```{r}
plot.lsmeans_bw.gr <- ggplot() + 
  #2
geom_point(data=force.out.greti_bal, aes(hr, bodyweight)) + 
  #3
  geom_point(data=greti_hr, aes(x=hr, y=fit), color="grey", alpha = 1/100) +
  #4
  geom_line(data=greti_hr, aes(x=hr, y=fit), color="blue") +
  #5
  geom_ribbon(data= greti_hr, aes(x=hr, ymin=lower, ymax=upper), alpha= 0.3, fill="blue")+
  #6
  xlab("hours across the day")+
  ylab("bodyweight (g) - greti")
  



plot.lsmeans0.gr <- 
  ggplot() +  
  geom_point(data = plot.acc0.gr, aes(x = age, y = lsmean, alpha = 1)) +
  geom_errorbar(data = plot.acc0.gr, aes(x = age, ymin = lower.CL, ymax = upper.CL, width = 0.5),alpha = 1) +
    ylab(expression(inital~leg~acceleration~(m/s^2)-greti)) +
  theme(axis.title.x=element_blank()) + #getting rid of x-axis title
  theme(legend.position = "none") 



#plot.lsmeans1.gr <- 
#  ggplot() +  
#  geom_point(data = plot.acc1.gr, aes(x = age, y = #lsmean), alpha = 1) +
#  geom_errorbar(data = plot.acc1.gr, aes(x = age, ymin = lower.CL, ymax = upper.CL, width = 0.5),alpha = 1) +
#  ylab(expression(acc~1~(m/s^2)-greti)) +
  #ylim(-2,6) +
#  theme(axis.title.x=element_blank()) + #getting rid of x-axis title
#  theme(legend.position = "none") 


#plot.lsmeans2.gr <- 
#  ggplot() +  
#  geom_point(data = plot.acc2.gr, aes(x = age, y = lsmean), alpha = 1) +
#  geom_errorbar(data = plot.acc2.gr, aes(x = age, ymin = lower.CL, ymax = upper.CL, width = 0.5),alpha = 1) +
#  ylab(expression(acc~2~(m/s^2)-greti)) +
  #ylim(-2,6) +
#  theme(axis.title.x=element_blank()) + #getting rid of x-axis title
#  theme(legend.position = "none") 



plot.lsmeans3.gr <- 
  ggplot() +  
  geom_point(data = plot.acc3.gr, aes(x = age, y = lsmean), alpha = 1) +
  geom_errorbar(data = plot.acc3.gr, aes(x = age, ymin = lower.CL, ymax = upper.CL, width = 0.5),alpha = 1) +
  ylab(expression(acc~3~(m/s^2)-greti)) +
  #ylim(-2,6) +
  theme(axis.title.x=element_blank()) + #getting rid of x-axis title
  theme(legend.position = "none") 

```
##### show plots  
```{r}
plot.lsmeans_bw.gr

grid.arrange(plot.lsmeans0.gr, plot.lsmeans3.gr, nrow = 1)
```
## H. raw graphs
### 1. plot data 
#### bluti
```{r}

#group by 

force.out.bluti.raw_bw <- force.out.bluti %>%
  group_by(day_since, group, scare_outcome,date_label) %>%
  summarise(
    mean = mean(bodyweight, na.rm = TRUE),
    SE = sd(bodyweight, na.rm = TRUE) / sqrt(n()),
    df = n() - 1,
    lower.CL = mean_cl_normal(bodyweight)$ymin,
    upper.CL = mean_cl_normal(bodyweight)$ymax
  ) 

force.out.bluti.raw_bw 

force.out.bluti.raw_acc0 <- force.out.bluti %>%
  group_by(day_since, group, scare_outcome,date_label) %>%
  summarise(
    mean = mean(int_acc, na.rm = TRUE),
    SE = sd(int_acc, na.rm = TRUE) / sqrt(n()),
    df = n() - 1,
    lower.CL = mean_cl_normal(int_acc)$ymin,
    upper.CL = mean_cl_normal(int_acc)$ymax
  ) 

force.out.bluti.raw_acc0

force.out.bluti.raw_maxFtot_scare <- force.out.bluti %>%
  group_by(scare_outcome) %>%
  summarise(
    mean = mean(max_Ftot, na.rm = TRUE),
    SE = sd(max_Ftot, na.rm = TRUE) / sqrt(n()),
    df = n() - 1,
    lower.CL = mean_cl_normal(max_Ftot)$ymin,
    upper.CL = mean_cl_normal(max_Ftot)$ymax
    
  )

force.out.bluti.raw_maxFtot_scare

force.out.bluti.raw_maxFtot_days <- force.out.bluti %>%
  group_by(day_since,group) %>%
  summarise(
    mean = mean(max_Ftot, na.rm = TRUE),
    SE = sd(max_Ftot, na.rm = TRUE) / sqrt(n()),
    df = n() - 1,
    lower.CL = mean_cl_normal(max_Ftot)$ymin,
    upper.CL = mean_cl_normal(max_Ftot)$ymax
    
  )

force.out.bluti.raw_maxFtot_days



force.out.bluti.raw_maxFx_scare <- force.out.bluti %>%
  group_by(scare_outcome) %>%
  summarise(
    mean = mean(Fx_maxFtot, na.rm = TRUE),
    SE = sd(Fx_maxFtot, na.rm = TRUE) / sqrt(n()),
    df = n() - 1,
    lower.CL = mean_cl_normal(Fx_maxFtot)$ymin,
    upper.CL = mean_cl_normal(Fx_maxFtot)$ymax
    
  )

force.out.bluti.raw_maxFx_scare

force.out.bluti.raw_maxFy_scare <- force.out.bluti %>%
  group_by(scare_outcome) %>%
  summarise(
    mean = mean(Fy_maxFtot, na.rm = TRUE),
    SE = sd(Fy_maxFtot, na.rm = TRUE) / sqrt(n()),
    df = n() - 1,
    lower.CL = mean_cl_normal(Fy_maxFtot)$ymin,
    upper.CL = mean_cl_normal(Fy_maxFtot)$ymax
    
  )

force.out.bluti.raw_maxFy_scare

force.out.bluti.raw_maxFx_days <- force.out.bluti %>%
  group_by(day_since,group) %>%
  summarise(
    mean = mean(Fx_maxFtot, na.rm = TRUE),
    SE = sd(Fx_maxFtot, na.rm = TRUE) / sqrt(n()),
    df = n() - 1,
    lower.CL = mean_cl_normal(Fx_maxFtot)$ymin,
    upper.CL = mean_cl_normal(Fx_maxFtot)$ymax
    
  )

force.out.bluti.raw_maxFx_days

force.out.bluti.raw_maxFy_days <- force.out.bluti %>%
  group_by(day_since,group) %>%
  summarise(
    mean = mean(Fy_maxFtot, na.rm = TRUE),
    SE = sd(Fy_maxFtot, na.rm = TRUE) / sqrt(n()),
    df = n() - 1,
    lower.CL = mean_cl_normal(Fy_maxFtot)$ymin,
    upper.CL = mean_cl_normal(Fy_maxFtot)$ymax
    
  )

force.out.bluti.raw_maxFy_days


ir.out.bluti.raw_acc1 <- ir.out.bluti %>%
  group_by(day_since, group, scare_outcome, date_label) %>%
  summarise(
    mean = mean(acc1, na.rm = TRUE),
    SE = sd(acc1, na.rm = TRUE) / sqrt(n()),
    df = n() - 1,
    lower.CL = mean_cl_normal(acc1)$ymin,
    upper.CL = mean_cl_normal(acc1)$ymax
  ) 

#ir.out.bluti.raw_acc1

ir.out.bluti.raw_acc2 <- ir.out.bluti %>%
  group_by(day_since, group, scare_outcome,date_label) %>%
  summarise(
    mean = mean(acc2, na.rm = TRUE),
    SE = sd(acc2, na.rm = TRUE) / sqrt(n()),
    df = n() - 1,
    lower.CL = mean_cl_normal(acc2)$ymin,
    upper.CL = mean_cl_normal(acc2)$ymax
  ) 

#ir.out.bluti.raw_acc2

ir.out.bluti.raw_acc3 <- ir.out.bluti %>%
  group_by(day_since, group, scare_outcome,date_label) %>%
  summarise(
    mean = mean(acc3, na.rm = TRUE),
    SE = sd(acc3, na.rm = TRUE) / sqrt(n()),
    df = n() - 1,
    lower.CL = mean_cl_normal(acc3)$ymin,
    upper.CL = mean_cl_normal(acc3)$ymax
  ) 
```

#### greti
```{r}
#group by 


force.out.greti.raw_bw <- force.out.greti %>%
  group_by(day_since, group, scare_outcome,date_label) %>%
  summarise(
    mean = mean(bodyweight, na.rm = TRUE),
    SE = sd(bodyweight, na.rm = TRUE) / sqrt(n()),
    df = n() - 1,
    lower.CL = mean_cl_normal(bodyweight)$ymin,
    upper.CL = mean_cl_normal(bodyweight)$ymax
  ) 

force.out.greti.raw_bw 


force.out.greti.raw_acc0 <- force.out.greti %>%
  group_by(day_since, group, scare_outcome,date_label) %>%
  summarise(
    mean = mean(int_acc, na.rm = TRUE),
    SE = sd(int_acc, na.rm = TRUE) / sqrt(n()),
    df = n() - 1,
    lower.CL = mean_cl_normal(int_acc)$ymin,
    upper.CL = mean_cl_normal(int_acc)$ymax
  ) 


force.out.greti.raw_acc0

force.out.greti.raw_maxFtot <- force.out.greti %>%
  group_by(day_since, group, scare_outcome, date_label) %>%
  summarise(
    mean = mean(max_Ftot, na.rm = TRUE),
    SE = sd(max_Ftot, na.rm = TRUE) / sqrt(n()),
    df = n() - 1,
    lower.CL = mean_cl_normal(max_Ftot)$ymin,
    upper.CL = mean_cl_normal(max_Ftot)$ymax
    
  )

force.out.greti.raw_maxFtot

force.out.greti.raw_maxFx <- force.out.greti %>%
  group_by(day_since, group, scare_outcome, date_label) %>%
  summarise(
    mean = mean(Fx_maxFtot, na.rm = TRUE),
    SE = sd(Fx_maxFtot, na.rm = TRUE) / sqrt(n()),
    df = n() - 1,
    lower.CL = mean_cl_normal(Fx_maxFtot)$ymin,
    upper.CL = mean_cl_normal(Fx_maxFtot)$ymax
    
  )

force.out.greti.raw_maxFy <- force.out.greti %>%
  group_by(day_since, group, scare_outcome, date_label) %>%
  summarise(
    mean = mean(Fy_maxFtot, na.rm = TRUE),
    SE = sd(Fy_maxFtot, na.rm = TRUE) / sqrt(n()),
    df = n() - 1,
    lower.CL = mean_cl_normal(Fy_maxFtot)$ymin,
    upper.CL = mean_cl_normal(Fy_maxFtot)$ymax
    
  )

force.out.greti.raw_maxFx_scare <- force.out.greti %>%
  group_by(scare_outcome) %>%
  summarise(
    mean = mean(Fx_maxFtot, na.rm = TRUE),
    SE = sd(Fx_maxFtot, na.rm = TRUE) / sqrt(n()),
    df = n() - 1,
    lower.CL = mean_cl_normal(Fx_maxFtot)$ymin,
    upper.CL = mean_cl_normal(Fx_maxFtot)$ymax
    
  )

force.out.greti.raw_maxFx_scare

force.out.greti.raw_maxFy_scare <- force.out.greti %>%
  group_by(scare_outcome) %>%
  summarise(
    mean = mean(Fy_maxFtot, na.rm = TRUE),
    SE = sd(Fy_maxFtot, na.rm = TRUE) / sqrt(n()),
    df = n() - 1,
    lower.CL = mean_cl_normal(Fy_maxFtot)$ymin,
    upper.CL = mean_cl_normal(Fy_maxFtot)$ymax
    
  )

force.out.greti.raw_maxFy_scare

force.out.greti.raw_maxFx_days <- force.out.greti %>%
  group_by(day_since) %>%
  summarise(
    mean = mean(Fx_maxFtot, na.rm = TRUE),
    SE = sd(Fx_maxFtot, na.rm = TRUE) / sqrt(n()),
    df = n() - 1,
    lower.CL = mean_cl_normal(Fx_maxFtot)$ymin,
    upper.CL = mean_cl_normal(Fx_maxFtot)$ymax
    
  )

force.out.greti.raw_maxFx_days

force.out.greti.raw_maxFy_days <- force.out.greti %>%
  group_by(day_since) %>%
  summarise(
    mean = mean(Fy_maxFtot, na.rm = TRUE),
    SE = sd(Fy_maxFtot, na.rm = TRUE) / sqrt(n()),
    df = n() - 1,
    lower.CL = mean_cl_normal(Fy_maxFtot)$ymin,
    upper.CL = mean_cl_normal(Fy_maxFtot)$ymax
    
  )

force.out.greti.raw_maxFy_days

ir.out.greti.raw_acc1 <- ir.out.greti %>%
  group_by(day_since, group, scare_outcome,date_label) %>%
  summarise(
    mean = mean(acc1, na.rm = TRUE),
    SE = sd(acc1, na.rm = TRUE) / sqrt(n()),
    df = n() - 1,
    lower.CL = mean_cl_normal(acc2)$ymin,
    upper.CL = mean_cl_normal(acc2)$ymax
  ) 


#ir.out.bluti.raw_acc1



ir.out.greti.raw_acc2 <- ir.out.greti %>%
  group_by(day_since, group, scare_outcome,date_label) %>%
  summarise(
    mean = mean(acc2, na.rm = TRUE),
    SE = sd(acc2, na.rm = TRUE) / sqrt(n()),
    df = n() - 1,
    lower.CL = mean_cl_normal(acc2)$ymin,
    upper.CL = mean_cl_normal(acc2)$ymax
  ) 

#ir.out.bluti.raw_acc2

ir.out.greti.raw_acc3 <- ir.out.greti %>%
  group_by(day_since, group, scare_outcome,date_label) %>%
  summarise(
    mean = mean(acc3, na.rm = TRUE),
    SE = sd(acc3, na.rm = TRUE) / sqrt(n()),
    df = n() - 1,
    lower.CL = mean_cl_normal(acc3)$ymin,
    upper.CL = mean_cl_normal(acc3)$ymax
  ) 
```

### 2. graphs - bluti
#### bodyweight
```{r}
plot_raw_bw_bluti <- 
  ggplot() +  
  geom_point(data = force.out.bluti.raw_bw , aes(x = day_since, y = mean, alpha = 1, color = day_since)) +
  geom_errorbar(data = force.out.bluti.raw_bw , aes(x = day_since, ymin = lower.CL, ymax = upper.CL, width = 0.5, color = day_since),alpha = 1, ) +
  ylab("bodyweight (g)-bluti") +
  #ylim(0.014,0.018) +
  theme(axis.title.x=element_blank()) + #getting rid of x-axis title
  theme(legend.position = "none") +
  facet_grid(date_label ~scare_outcome+ group)

plot_raw_bw_bluti

```
#### initial acceleration (acc0)
```{r}
plot_raw_acc0_bluti <- 
  ggplot() +  
  geom_point(data = force.out.bluti.raw_acc0 , aes(x = day_since, y = mean, alpha = 1, color = day_since)) +
  geom_errorbar(data = force.out.bluti.raw_acc0 , aes(x = day_since, ymin = lower.CL, ymax = upper.CL, width = 0.5, color = day_since),alpha = 1, ) +
  ylab(expression(initial~acceleration~(m/s^2)-bluti)) +
  #ylim(0.014,0.018) +
  theme(axis.title.x=element_blank()) + #getting rid of x-axis title
  theme(legend.position = "none") +
  facet_grid(date_label ~scare_outcome+ group)

plot_raw_acc0_bluti

#Jan21+1 data have to reverse engineer it. 
```

#### maxFtot
##### normal vs. escape
```{r}
plot_raw_maxFtot_bluti_scare <- 
  ggplot() +  
  geom_point(data = force.out.bluti.raw_maxFtot_scare , aes(x = as.factor(scare_outcome), y = mean), alpha = 1) +
  geom_errorbar(data = force.out.bluti.raw_maxFtot_scare , aes(x = as.factor(scare_outcome), ymin = lower.CL, ymax = upper.CL, width = 0.5),alpha = 1 ) +
  ylab("peak resultant force (N)") +
  scale_x_discrete(labels = c("normal", "escape"))+
  #ylim(0.014,0.018) +
  theme(axis.title.x=element_blank()) + #getting rid of x-axis title
  theme(legend.position = "none") #+
  #facet_grid(date_label ~scare_outcome+ group)

plot_raw_maxFtot_bluti_scare

```
##### day_since
```{r}
plot_raw_maxFtot_bluti_days <- 
  ggplot() +  
  geom_point(data = force.out.bluti.raw_maxFtot_days , aes(x = as.factor(day_since), y = mean), alpha = 1) +
  geom_errorbar(data = force.out.bluti.raw_maxFtot_days , aes(x = as.factor(day_since), ymin = lower.CL, ymax = upper.CL, width = 0.5),alpha = 1 ) +
  ylab("peak resultant force (N)") +
  xlab("day since experimental day")+
  #ylim(0.014,0.018) +
  #theme(axis.title.x=element_blank()) + #getting rid of x-axis title
  theme(legend.position = "none") +
  facet_grid(~group)

plot_raw_maxFtot_bluti_days

#Jan21+1 data have to reverse engineer it. 
```

#### maxFx
##### normal vs. escape take-off
```{r}
plot_raw_maxFx_bluti_scare <- 
  ggplot() +  
  geom_point(data = force.out.bluti.raw_maxFx_scare , aes(x = as.factor(scare_outcome), y = mean, alpha = 1)) +
  geom_errorbar(data = force.out.bluti.raw_maxFx_scare , aes(x = as.factor(scare_outcome), ymin = lower.CL, ymax = upper.CL, width = 0.5),alpha = 1 ) +
  ylab("vertical force (N)") +
  #ylim(0.014,0.018) +
  scale_x_discrete(labels = c('normal','escape'))+
  theme(axis.title.x=element_blank()) + #getting rid of x-axis title
  theme(legend.position = "none") #+
  #facet_grid(date_label ~scare_outcome+ group)

plot_raw_maxFx_bluti_scare

#Jan21+1 data have to reverse engineer it. 
```
##### day_since
```{r}
plot_raw_maxFx_bluti_days <- 
  ggplot() +  
  geom_point(data = force.out.bluti.raw_maxFx_days , aes(x = as.factor(day_since), y = mean, alpha = 1)) +
  geom_errorbar(data = force.out.bluti.raw_maxFx_days , aes(x = as.factor(day_since), ymin = lower.CL, ymax = upper.CL, width = 0.5),alpha = 1, ) +
  ylab("vertical force (N)") +
  xlab("day since experimental day")+
  #ylim(0.014,0.018) +
  #theme(axis.title.x=element_blank()) + #getting rid of x-axis title
  theme(legend.position = "none") +
  facet_grid(~group)

plot_raw_maxFx_bluti_days
```

#### maxFy
##### normal vs. escape take-off
```{r}
plot_raw_maxFy_bluti_scare <- 
  ggplot() +  
  geom_point(data = force.out.bluti.raw_maxFy_scare , aes(x = as.factor(scare_outcome), y = -mean, alpha = 1)) +
  geom_errorbar(data = force.out.bluti.raw_maxFy_scare , aes(x = as.factor(scare_outcome), ymin = -lower.CL, ymax = -upper.CL, width = 0.5),alpha = 1, ) +
  ylab("horizontal force (N)") +
  #ylim(0.014,0.018) +
  scale_x_discrete(labels = c('normal','escape'))+
  theme(axis.title.x=element_blank()) + #getting rid of x-axis title
  theme(legend.position = "none") #+
  #facet_grid(date_label ~scare_outcome+ group)

plot_raw_maxFy_bluti_scare

#Jan21+1 data have to reverse engineer it. 
```
##### day_since
```{r}
plot_raw_maxFy_bluti_days <- 
  ggplot() +  
  geom_point(data = force.out.bluti.raw_maxFy_days , aes(x = as.factor(day_since), y = -mean, alpha = 1)) +
  geom_errorbar(data = force.out.bluti.raw_maxFy_days , aes(x = as.factor(day_since), ymin = -lower.CL, ymax = -upper.CL, width = 0.5),alpha = 1, ) +
  ylab("horizontal force (N)") +
  xlab("day since experimental day")+
  #ylim(0.014,0.018) +
  #theme(axis.title.x=element_blank()) + #getting rid of x-axis title
  theme(legend.position = "none") +
  facet_grid( ~group)

plot_raw_maxFy_bluti_days
```
##### collaged
```{r}
grid.arrange(plot_raw_maxFtot_bluti_days, plot_raw_maxFtot_bluti_scare, plot_raw_maxFx_bluti_days, plot_raw_maxFx_bluti_scare, plot_raw_maxFy_bluti_days, plot_raw_maxFy_bluti_scare, nrow = 3, top = "Raw blue tit leg take-off force")
```

#### acc1 
```{r}
plot_raw_acc1_bluti <- 
  ggplot() +  
  geom_point(data = ir.out.bluti.raw_acc1 , aes(x = day_since, y = mean, alpha = 1, color = day_since)) +
  geom_errorbar(data = ir.out.bluti.raw_acc1 , aes(x = day_since, ymin = lower.CL, ymax = upper.CL, width = 0.5, color = day_since),alpha = 1, ) +
  ylab(expression(acc1~(m/s^2)-bluti)) +
  #ylim(0.014,0.018) +
  theme(axis.title.x=element_blank()) + #getting rid of x-axis title
  theme(legend.position = "none") +
  facet_grid(date_label ~scare_outcome+ group)

plot_raw_acc1_bluti
```

#### acc2
```{r}
plot_raw_acc2_bluti <- 
  ggplot() +  
  geom_point(data = ir.out.bluti.raw_acc2 , aes(x = day_since, y = mean, alpha = 1, color = day_since)) +
  geom_errorbar(data = ir.out.bluti.raw_acc2 , aes(x = day_since, ymin = lower.CL, ymax = upper.CL, width = 0.5, color = day_since),alpha = 1, ) +
  ylab(expression(acc2~(m/s^2)-bluti)) +
  #ylim(0.014,0.018) +
  theme(axis.title.x=element_blank()) + #getting rid of x-axis title
  theme(legend.position = "none") +
  facet_grid(date_label ~scare_outcome+ group)

plot_raw_acc2_bluti
```
#### acc3
```{r}
plot_raw_acc3_bluti <- 
  ggplot() +  
  geom_point(data = ir.out.bluti.raw_acc3 , aes(x = day_since, y = mean, alpha = 1, color = day_since)) +
  geom_errorbar(data = ir.out.bluti.raw_acc3 , aes(x = day_since, ymin = lower.CL, ymax = upper.CL, width = 0.5, color = day_since),alpha = 1, ) +
  ylab(expression(acc3~(m/s^2)-bluti)) +
  #ylim(0.014,0.018) +
  theme(axis.title.x=element_blank()) + #getting rid of x-axis title
  theme(legend.position = "none") +
  facet_grid(date_label ~scare_outcome+ group)

plot_raw_acc3_bluti
```
### 3.graphs - greti
#### bodyweight - greti
```{r}
plot_raw_bw_greti <- 
  ggplot() +  
  geom_point(data = force.out.greti.raw_bw , aes(x = day_since, y = mean, alpha = 1, color = day_since)) +
  geom_errorbar(data = force.out.greti.raw_bw , aes(x = day_since, ymin = lower.CL, ymax = upper.CL, width = 0.5, color = day_since),alpha = 1, ) +
  ylab("bodyweight (g)-greti") +
  #ylim(0.014,0.018) +
  theme(axis.title.x=element_blank()) + #getting rid of x-axis title
  theme(legend.position = "none") +
  facet_grid(date_label ~scare_outcome+ group)

plot_raw_bw_greti

```
#### initial acceleration (acc0) - greti
```{r}
plot_raw_acc0_greti <- 
  ggplot() +  
  geom_point(data = force.out.greti.raw_acc0 , aes(x = day_since, y = mean, alpha = 1, color = day_since)) +
  geom_errorbar(data = force.out.greti.raw_acc0 , aes(x = day_since, ymin = lower.CL, ymax = upper.CL, width = 0.5, color = day_since),alpha = 1, ) +
  ylab(expression(initial~acceleration~(m/s^2)-greti)) +
  #ylim(0.014,0.018) +
  theme(axis.title.x=element_blank()) + #getting rid of x-axis title
  theme(legend.position = "none") +
  facet_grid(date_label ~scare_outcome+ group)

plot_raw_acc0_greti

#Jan21+1 data have to reverse engineer it. 
```

#### maxFtot
```{r}
plot_raw_maxFtot_greti <- 
  ggplot() +  
  geom_point(data = force.out.greti.raw_maxFtot , aes(x = day_since, y = mean, alpha = 1, color = day_since)) +
  geom_errorbar(data = force.out.greti.raw_maxFtot , aes(x = day_since, ymin = lower.CL, ymax = upper.CL, width = 0.5, color = day_since),alpha = 1, ) +
  ylab("max total leg force (N)-greti") +
  #ylim(0.014,0.018) +
  theme(axis.title.x=element_blank()) + #getting rid of x-axis title
  theme(legend.position = "none") +
  facet_grid(date_label ~scare_outcome+ group)

plot_raw_maxFtot_greti

#Jan21+1 data have to reverse engineer it. 
```

#### maxFx
```{r}
plot_raw_maxFx_greti <- 
  ggplot() +  
  geom_point(data = force.out.greti.raw_maxFx , aes(x = day_since, y = mean, alpha = 1, color = day_since)) +
  geom_errorbar(data = force.out.greti.raw_maxFx , aes(x = day_since, ymin = lower.CL, ymax = upper.CL, width = 0.5, color = day_since),alpha = 1, ) +
  ylab("max vertical leg force (N)-greti") +
  #ylim(0.014,0.018) +
  theme(axis.title.x=element_blank()) + #getting rid of x-axis title
  theme(legend.position = "none") +
  facet_grid(date_label ~scare_outcome+ group)

plot_raw_maxFx_greti

#Jan21+1 data have to reverse engineer it. 
```

#### maxFy
```{r}
plot_raw_maxFy_greti <- 
  ggplot() +  
  geom_point(data = force.out.greti.raw_maxFy , aes(x = day_since, y = mean, alpha = 1, color = day_since)) +
  geom_errorbar(data = force.out.greti.raw_maxFy , aes(x = day_since, ymin = lower.CL, ymax = upper.CL, width = 0.5, color = day_since),alpha = 1, ) +
  ylab("max horizontal leg force (N)-greti") +
  #ylim(0.014,0.018) +
  theme(axis.title.x=element_blank()) + #getting rid of x-axis title
  theme(legend.position = "none") +
  facet_grid(date_label ~scare_outcome+ group)

plot_raw_maxFy_greti

#Jan21+1 data have to reverse engineer it. 
```

#### acc1 
```{r}
plot_raw_acc1_greti <- 
  ggplot() +  
  geom_point(data = ir.out.greti.raw_acc1 , aes(x = day_since, y = mean, alpha = 1, color = day_since)) +
  geom_errorbar(data = ir.out.greti.raw_acc1 , aes(x = day_since, ymin = lower.CL, ymax = upper.CL, width = 0.5, color = day_since),alpha = 1, ) +
  ylab(expression(acc1~(m/s^2)-greti)) +
  #ylim(0.014,0.018) +
  theme(axis.title.x=element_blank()) + #getting rid of x-axis title
  theme(legend.position = "none") +
  facet_grid(date_label ~scare_outcome+ group)

plot_raw_acc1_greti
```

#### acc2
```{r}
plot_raw_acc2_greti <- 
  ggplot() +  
  geom_point(data = ir.out.greti.raw_acc2 , aes(x = day_since, y = mean, alpha = 1, color = day_since)) +
  geom_errorbar(data = ir.out.greti.raw_acc2 , aes(x = day_since, ymin = lower.CL, ymax = upper.CL, width = 0.5, color = day_since),alpha = 1, ) +
  ylab(expression(acc2~(m/s^2)-greti)) +
  #ylim(0.014,0.018) +
  theme(axis.title.x=element_blank()) + #getting rid of x-axis title
  theme(legend.position = "none") +
  facet_grid(date_label ~scare_outcome+ group)

plot_raw_acc2_greti
```
#### acc3
```{r}
plot_raw_acc3_greti <- 
  ggplot() +  
  geom_point(data = ir.out.greti.raw_acc3 , aes(x = day_since, y = mean, alpha = 1, color = day_since)) +
  geom_errorbar(data = ir.out.greti.raw_acc3 , aes(x = day_since, ymin = lower.CL, ymax = upper.CL, width = 0.5, color = day_since),alpha = 1, ) +
  ylab(expression(acc3~(m/s^2)-greti)) +
  #ylim(0.014,0.018) +
  theme(axis.title.x=element_blank()) + #getting rid of x-axis title
  theme(legend.position = "none") +
  facet_grid(date_label ~scare_outcome+ group)

plot_raw_acc3_greti
```
#### takeoff-angle (optional) {.tabset}
##### model assumptions
-takeoff_angle probably not good to test -> deviation significant
```{r}
lmer_angle_bluti = lmer(takeoff_angle ~ 1 + day_since*group  +scare_outcome + hr + sex + age+ celsius.now + flight_no + date +(1|id), data = force.out.bluti_bal)


# Computes the default residuals (raw conditional)
raw_cond_angle <- compute_redres(lmer_angle_bluti)

#linearity - can't test as all the terms are categorical 
hist(raw_cond_angle) #normality in histogram
plot_redres(lmer_angle_bluti) #heteroskedasticity
plot_resqq(lmer_angle_bluti) #normality of residuals - heavy tailed and bimodal?
plot_ranef(lmer_angle_bluti) #normality of random effects residuals

simulation_output <- simulateResiduals(fittedModel = lmer_angle_bluti, n = 1000)

plot(simulation_output)

```

##### model analysis
###### lm4e
```{r}
#Anova(lmer_bw_bluti, type = 3, test.statistic = "F") 
#emm_options(pbkrtest.limit = 7000)
tab_model(lmer_angle_bluti)
lsmeans_angle_bluti <- lsmeans(lmer_angle_bluti, pairwise~day_since*group)

#,lmerTest.limit = 7000
#,lmerTest.limit = 7000
```
###### CRVE
```{r}

#force.out.bluti_bal <- force.out.bluti_bal %>%
#  rename(group = exp_group)
plm_angle_bluti_random <- plm(takeoff_angle ~ 0+ day_since*group  +scare_outcome + hr + sex + age+ celsius.now + flight_no, data = force.out.bluti_bal, effect = "individual", index = c("id", "visit_datetime"), model = "random")

plm_angle_bluti_fixed <- plm(takeoff_angle ~ 0+ day_since*group  +scare_outcome + sex + age+ celsius.now + flight_no, data = force.out.bluti_bal, effect = "individual", index = c("id", "visit_datetime"), model = "within")

#Hausman test
hausman_test <- phtest(plm_angle_bluti_random, plm_angle_bluti_fixed)
print(hausman_test)

#Fixed effects model:
# Controls for time-invariant individual heterogeneity by allowing intercepts to vary across entities.
#Random effects model:
#Assumes that individual heterogeneity is random and uncorrelated with the explanatory variables.
#p<0.05, fixed model preferred. 
```

##### group -> exp_group
```{r}
force.out.bluti_bal <- force.out.bluti_bal %>% rename(exp_group = group)
force.out.greti_bal <- force.out.greti_bal %>% rename(exp_group = group)

ir.out.bluti_bal <- ir.out.bluti_bal %>% rename(exp_group = group)
ir.out.greti_bal <- ir.out.greti_bal %>% rename(exp_group = group)

```
##### plm
```{r}
#run once
```
```{r}

#plm

plm_bw_bluti_fixed <- plm(int_acc ~ 0 + day_since*exp_group + scare_outcome + prev_scares + flight_no  + sex + age + hr+ celsius.now, data = force.out.bluti_bal, effect = "individual", index = c("id","date"), model = "within")

plot_model(plm_bw_bluti_fixed)






plot_predictions(plm_bw_bluti_fixed, by = c("day_since","exp_group"))
```
##### lmer
```{r}
emmeans_bw_bluti <- emmeans(lmer_bw_bluti, ~ exp_group | day_since, type = "response", level = 0.95, pbkrtest.limit = 5734) 

emmeans_bw_bluti_df <- as.data.frame(emmeans_bw_bluti)

#emmeans_bw_plm_df
emmeans_bw_bluti_df
#plot_prediction has to change "group" 

```
##### Perform pairwise comparisons of Group within each Object
```{r}
pairwise_contrast_bw <- contrast(emmeans_bw_bluti, method = "pairwise")

pairwise_contrast_bw <- contrast(emmeans_bw_bluti, method = "pairwise")

# Summarize contrasts and include confidence intervals (set infer = c(TRUE, TRUE))
contrast_df_bw3 <- summary(pairwise_contrast_bw, infer = c(TRUE, TRUE)) |> as.data.frame()

# View the resulting contrast estimates with associated uncertainty
contrast_df_bw3

```
##### CRVE
```{r}
#plm_bw_bluti_random <- plm(bodyweight ~ 0 + #day_since*group + scare_outcome + prev_scares #+ flight_no  + sex + age + hr+ date+ #celsius.now , data = force.out.bluti_bal, #effect = "individual", index = c("id"), model #= "random") 


```




